# ---
# title: "Preliminary SEM Analyses 2"
# author: "Delta Food Web Team"
# date: "10/25/2021"
# output: html_document
# knit: (function(input, ...) {
#     rmarkdown::render(
#       input,
#       output_dir = '../docs',
#       knit_root_dir = '..',
#       envir = globalenv()
#     )
#   })
# ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(dplyr)
library(psych) #for pairs.panels, but could use other packages, e.g. GGalley
library(lavaan)
# library(semPlot)
```

### Import data

```{r dataprep}
combined=read.csv("data/annual_averages/annual_data_compiled.csv")
cnames=read.csv("analysis/column_names.csv", stringsAsFactors = F)
dsub=filter(combined, Year>=1975)
focaldata=dsub[,cnames$Datacolumn]
fvars=cnames$Shortname
colnames(focaldata)=fvars
```

### Data prep

Log transform, scale

```{r histo}
#log transform
logvars=fvars[cnames$Log=="yes"]
logtrans=function(x) {
  x2=x[which(!is.na(x))]
  if(any(x2==0)) {log(x+min(x2[which(x2>0)],na.rm=T))}
  else {log(x)}
}
focaldatalog = focaldata %>% 
  mutate_at(logvars,logtrans)

#scale data
fd=focaldatalog
fd=fd %>% 
  mutate_at(2:length(fvars),list("1"=lag)) %>% #lag 1
  mutate_at(2:length(fvars),list("fd"=function(x) c(NA,diff(x)))) %>% #first difference
  mutate_at(2:length(fvars),list("dtr"=function(x) { #detrend
    x2=x
    x2[x2==0]=NA
    res=residuals(lm(x2~fd$year))
    out=x
    out[which(!is.na(x2))]=res
    return(out)
  })) %>%
  mutate_at(-1,scale)
```

### Time series plots

Original units

```{r timeseries1, fig.width=8, fig.height=11, echo=FALSE}
#original units
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=focaldata[,1],y=focaldata[,i],main=fvars[i], type="o",cex=0.5)
}
```

Log scaled

```{r timeseries2, fig.width=8, fig.height=11, echo=FALSE}
#log scaled
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i],main=fvars[i], type="o",cex=0.5)
}
```

First difference

```{r timeseries3, fig.width=8, fig.height=11, echo=FALSE}
#first diff
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i+2*(length(fvars)-1)],main=fvars[i], type="o",cex=0.5)
}
```

Detrended

```{r timeseries4, fig.width=8, fig.height=11, echo=FALSE}
#detrended
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i+3*(length(fvars)-1)],main=fvars[i], type="o",cex=0.5)
}
```

### Bivariate plots

```{r bivariate, fig.width=9, fig.height=9, echo=FALSE}
psych::pairs.panels(dplyr::select(fd, year, flow, temp, secchi, chla, hcope, clad, pcope, mysid, estfish, sside, clams, tphos),lm = T)
```

### SEM model

Best model so far

```{r}
model4='zoop=~hcope+clad+mysid
        fish=~estfish+estfish_bsmt+estfish_bsot
        zoop~sside+chla
        chla~clams+flow
        fish~zoop+flow+sside
        hcope~~mysid
        clad~~mysid
        hcope~~clad'
modfit4=sem(model4, data=fd)
summary(modfit4, standardized=T, rsq=T)
semPaths(modfit4, "std", edge.label.cex = 1, residuals = F)
semPaths(modfit4, "par", edge.label.cex = 1, residuals = F)
```

Same model with detrended data

```{r}
model4b='zoop=~hcope_dtr+clad_dtr+mysid_dtr
        fish=~estfish_dtr+estfish_bsmt_dtr+estfish_bsot_dtr
        zoop~sside_dtr+chla_dtr
        chla_dtr~clams_dtr+flow_dtr
        fish~zoop+flow_dtr+sside_dtr'
modfit4b=sem(model4b, data=fd)
summary(modfit4b, standardized=T, rsq=T)
semPaths(modfit4b, "std", edge.label.cex = 1, residuals = F)
semPaths(modfit4b, "par", edge.label.cex = 1, residuals = F)
```

Old models (not run)

```{r, eval=FALSE}
model1='clams~year+flow
        sside~year+flow
        chla~flow+clams #+year?
        hcope~chla+pcope #+year?
        clad~chla+pcope
        #pcope~year+clams
        mysid~hcope+pcope
        estfish~hcope+pcope+mysid+sside'
modfit1=sem(model1, data=fd)
summary(modfit1, standardized=T, rsq=T)
semPaths(modfit1, "std", edge.label.cex = 1, residuals = F)

residuals(modfit1)

model1='clams~year+flow
        sside~year+flow
        foodsupply=~chla+hcope+clad+mysid
        foodsupply~clams+flow+sside
        fish=~estfish+estfish_bsmt
        fish~foodsupply'
modfit1=sem(model1, data=fd)
summary(modfit1, standardized=T, rsq=T)
semPaths(modfit1, "std", edge.label.cex = 1, residuals = F)

model1='foodsupply=~chla+hcope+clad+mysid
        fish=~estfish+estfish_bsmt
        foodsupply~clams+flow+sside
        fish~foodsupply'
modfit1=sem(model1, data=fd)
summary(modfit1, standardized=T, rsq=T)
semPaths(modfit1, "std", edge.label.cex = 1, residuals = F)

model2='foodsupply=~chla+hcope+clad+mysid
        fish=~estfish+estfish_bsmt
        foodsupply~clams+flow+sside
        fish~foodsupply+flow+sside'
modfit2=sem(model2, data=fd)
summary(modfit2, standardized=T, rsq=T)
semPaths(modfit2, "std", edge.label.cex = 1, residuals = F)

residuals(modfit2)


model3='zoop=~hcope+clad+mysid
        fish=~estfish+estfish_bsmt
        zoop~clams+flow+sside+chla
        chla~clams #+tphos+nitrate+ammonia
        fish~zoop+flow+sside'
modfit3=sem(model3, data=fd)
summary(modfit3, standardized=T, rsq=T)
semPaths(modfit3, "std", edge.label.cex = 1, residuals = F)

residuals(modfit3, type="cor")
modificationindices(modfit3)

model4='zoop=~hcope+clad+mysid
        fish=~estfish+estfish_bsmt+estfish_bsot
        zoop~sside+chla
        chla~clams+flow
        fish~zoop+flow+sside
        hcope~~mysid
        clad~~mysid
        hcope~~clad'
modfit4=sem(model4, data=fd)
summary(modfit4, standardized=T, rsq=T)
semPaths(modfit4, "std", edge.label.cex = 1, residuals = F)
semPaths(modfit4, "par", edge.label.cex = 1, residuals = F)

residuals(modfit4, type="cor")
modificationindices(modfit4)
```

```{r, eval=FALSE}
#using estfish instead of smelt
model1f='chla~year+flow+secchi
        hcope~year+flow+temp+chla+pcope+secchi
        pcope~year
        mysid~year+hcope+pcope
        estfish~year+secchi+hcope+pcope+mysid'
modfit1f=sem(model1f, data=fd)
#summary(modfit1f, standardized=T, rsq=T)
semPaths(modfit1f, "std", edge.label.cex = 1, residuals = F)

#using marfish instead of smelt
model1g='chla~year+flow+secchi
        hcope~year+flow+temp+chla+pcope+secchi
        pcope~year
        mysid~year+hcope+pcope
        marfish~year+secchi+hcope+pcope+mysid'
modfit1g=sem(model1g, data=fd)
#summary(modfit1g, standardized=T, rsq=T)
semPaths(modfit1g, "std", edge.label.cex = 1, residuals = F)
```

Detrended (anomalies) without year effect

```{r, eval=FALSE}
#smelt
model2a='chla_dtr~flow_dtr+temp_dtr+secchi_dtr
         hzoop_dtr~flow_dtr+temp_dtr+secchi_dtr+chla_dtr+pcope_dtr
         pcope_dtr~flow_dtr+temp_dtr+secchi_dtr
         mysid_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr
         smelt_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr+mysid_dtr'
modfit2a=sem(model2a, data=fd)
#summary(modfit2a, standardized=T, rsq=T)
semPaths(modfit2a, "std", edge.label.cex = 1, residuals = F)

#estuarine fishes
model2b='chla_dtr~flow_dtr+temp_dtr+secchi_dtr
         hzoop_dtr~flow_dtr+temp_dtr+secchi_dtr+chla_dtr+pcope_dtr
         pcope_dtr~flow_dtr+temp_dtr+secchi_dtr
         mysid_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr
         estfish_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr+mysid_dtr'
modfit2b=sem(model2b, data=fd)
#summary(modfit2b, standardized=T, rsq=T)
semPaths(modfit2b, "std", edge.label.cex = 1, residuals = F)
```

