---
title: "Benthic Sampling Effort"
author: "Jeanette Clark"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(leaflet)
library(sf)
```


## Read and preprocess data

* remove the L,R,C (left, right, center) from site names, so we can group those counts together
    - may need to verify this doesn't create any statistical artifacts
* create a monthly time stamp
* take mean of lat/lon for each site
    -  some sites moved slightly, and some years don't have lat/lons for the sites. Assume minor site movement is negligable (may need to verify)

```{r, message=FALSE}
ben <- read_csv("data_raw/DWR Benthic raw count data 1975-2020 2021_10_01.csv")  %>% 
  mutate(StationClean =  gsub("-[A-Z]","",StationCode)) %>% 
  mutate(mon = lubridate::ym(paste0(Year, Month, sep = "-"))) %>% 
  group_by(StationClean) %>% 
  mutate(Latitude = mean(Latitude, na.rm = T), 
         Longitude = mean(Longitude, na.rm = T)) %>% 
  ungroup()
```



```{r}
sites_mon <- ben  %>% 
  distinct(StationClean, Latitude, Longitude, mon) 

sites <- ben %>% 
  group_by(StationClean, Latitude, Longitude) %>% 
  summarise(n_months = length(unique(mon)), .groups  =  "drop") %>% 
  st_as_sf(coords = c('Longitude','Latitude'), crs = 4326,  remove = FALSE)
```


## Sites that were sampled each month


```{r, echo=F}
ggplot(sites_mon, aes(x = mon, y = StationClean), size = 3) +
  geom_point() +
  theme_classic()
```


## Table of site information

```{r, echo = FALSE}
datatable(as.data.frame(sites))
```

## Core Stations

```{r}
core <- c("P8",
          "MD10",
          "D8",
          "D7",
          "D6",
          "D41",
          "D4",
          "D28A",
          "D26",
          "C3",
          "C10")

core_data <-  sites_mon %>% 
  filter(StationClean %in% core)
```


```{r, echo=F}
ggplot(core_data, aes(x = mon, y = StationClean), size = 3) +
  geom_point() +
  theme_classic()
```

## Location of sites, sized according to total number of months sampled

Red are core sites

```{r, message = FALSE, echo = FALSE}
sites_core <- sites %>% 
  filter(StationClean %in% core)


leaflet(sites) %>% 
  addTiles() %>% 
  addCircleMarkers(data = sites,
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = ~(n_months/50), # arbitrary scaling
                         fillColor = "blue",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~StationClean) %>% 
  addCircleMarkers(data = sites_core,
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = ~(n_months/50), # arbitrary scaling
                         fillColor = "red",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~StationClean)
```
