---
title: "dayflow yearly average calculations"
author: "Parsa Saffarinia"
date: "10/29/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
library(scales)
```


## read in data from website. 

* the dayflow is separated in year chunks so we have to download each chunk and stitch it together

    - may need to verify this doesn't create any statistical artifacts


```{r}
#Dayflow Results 1970 - 1983#
data1 <- read.csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/a0a46a1d-bec5-4db9-b331-655e306860ba/download/dayflow-results-1970-1983.csv")
#Dayflow Results 1984 - 1996
data2 <- read.csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/cb04e626-9729-4105-af81-f6e5a37f116a/download/dayflow-results-1984-1996.csv")
#Dayflow Results 1997 - 2020
data3 <- read.csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/21c377fe-53b8-4bd6-9e1f-2025221be095/download/dayflow-results-1997-2020.csv")
```



```{r}
sites_mon <- ben  %>% 
  distinct(StationClean, Latitude, Longitude, mon) 

sites <- ben %>% 
  group_by(StationClean, Latitude, Longitude) %>% 
  summarise(n_months = length(unique(mon)), .groups  =  "drop") %>% 
  st_as_sf(coords = c('Longitude','Latitude'), crs = 4326,  remove = FALSE)
```


## some nit-picky stuff, one column was spelled differently for one of the datasets (exports)
## then combine dataframes for each year chunk


```{r}
#fix "export" and "exports" that changed column names in different years
head(data1)
head(data2)
head(data3)
library(data.table)
setnames(data3, "EXPORTS", "EXPORT")

combo_dayflow = bind_rows(data1,data2, data3)
```


# converting calendar year to water year values and adding water year column
#months 10, 11, 12 = (calendar yr + 1). months 1-9, calendar yr = water yr

```{r}

w.year <- combo_dayflow$Year
oct.nov.dec <- (combo_dayflow$Month) > 9
w.year[oct.nov.dec] <- w.year[oct.nov.dec] + 1
combo_dayflow$Water_year <- w.year

#remove 1969 because there's no matching year it in water year, since sampling started in october.
combo_dayflow_CY<-combo_dayflow[-c(1:92), ]
```

## average total outflow, inflow, Yolo, Sac, export by calendar year and water year, rename columns to match

```{r}
##average total outflow, inflow, Yolo, Sac, export by calendar year
annual_average_deltawide_flow<-combo_dayflow_CY %>%
  group_by(Year) %>%
  summarise(value = mean(OUT), mean(TOT), mean(YOLO), mean(SAC), mean(EXPORT))
#name column for flow
annual_average_deltawide_flow_CY<-rename(annual_average_deltawide_flow,
c("Calendar_Year" = "Year","Mean_outflow_CY" = "value", "Mean_inflow_CY" = "mean(TOT)",
  "Mean_YOLO_CY" = "mean(YOLO)", "Mean_SAC_CY" = "mean(SAC)", "Mean_Export_CY" = "mean(EXPORT)"))

##average total outflow, inflow, Yolo, Sac, export by water year
WY_annual_average_deltawide_flow<-combo_dayflow %>%
  group_by(Water_year) %>%
  summarise(value = mean(OUT), mean(TOT), mean(YOLO), mean(SAC), mean(EXPORT))
#name column for flow
annual_average_deltawide_flow_WY<-
  rename(WY_annual_average_deltawide_flow,c("Mean_outflow_WY" = "value",
                                            "Mean_inflow_WY" = "mean(TOT)",

                                             "Mean_YOLO_WY" = "mean(YOLO)",
                                            "Mean_SAC_WY" = "mean(SAC)",
                                            "Mean_Export_WY" = "mean(EXPORT)"))
```

#combine the water year and calendar year metrics dataframes, write a CSV for analysis

```{r}
cy_wy_average_flow<- cbind(annual_average_deltawide_flow_CY,annual_average_deltawide_flow_WY)

library(readr)

write_csv(cy_wy_average_flow, file.path("annual_averages","annual_average_deltawide_flow.csv"))
```

