---
title: "Mississippi Silverside data from Delta Juvenile Fish Monitoring Program"
author: "Brian Mahardja"
date: "11/3/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
#devtools::install_github("sbashevkin/LTMRdata")

library(LTMRdata)
library(tidyverse)
library(lubridate)
library(ggmap)
source('../functions/nutrient_funcs.R', local = TRUE)

```

## Delta Juvenile Fish Monitoring Program

Interagency Ecological Program (IEP), R. McKenzie, J. Speegle, A. Nanninga, J.R. Cook, J. Hagen, and B. Mahardja. 2021. Interagency Ecological Program: Over four decades of juvenile fish monitoring data from the San Francisco Estuary, collected by the Delta Juvenile Fish Monitoring Program, 1976-2020 ver 7. Environmental Data Initiative. https://doi.org/10.6073/pasta/3d2b756ae45f43ddb895dc118f28b314

The United States Fish and Wildlife Service Delta Juvenile Fish Monitoring Program (DJFMP) has monitored juvenile Chinook Salmon Oncorhynchus tshawytscha and other fish species within the San Francisco Estuary (Estuary) since 1976 using a combination of surface trawls and beach seines. Since 2000, three trawl sites and 58 beach seine sites have been sampled weekly or biweekly within the Estuary and lower Sacramento and San Joaquin Rivers. As part of the Interagency Ecological Program (IEP) that manages the Estuary, the DJFMP has tracked the relative abundance and distribution of naturally and hatchery produced juvenile Chinook Salmon of all races as they outmigrate through the Sacramento-San Joaquin Delta for over four decades. The data that DJFMP collected has been used not only to help inform the management of Chinook Salmon, but also to monitor the status of native species of interest such as the previously listed Sacramento Splittail Pogonichthys macrolepidotus and invasive species such as Mississippi Silverside Menidia beryllina and Largemouth Bass Micropterus salmoides. For more information: https://www.fws.gov/lodi/juvenile_fish_monitoring_program/jfmp_index.htm  

## Load data and show sampling effort

Load data using Sam's LTMRdata package and show map of beach seine stations. 

```{r, echo = FALSE, message=FALSE, warning=FALSE}

DJFMP_Silverside_seine <-LTMRdata::fish(sources="DJFMP", species="Menidia audens", size_cutoff=NULL,remove_unknown_lengths=FALSE) %>%
  filter(Method=="Beach seine")

#Map
DJFMP_Coords<-DJFMP_Silverside_seine %>% group_by(Station) %>% summarise(Latitude=mean(Latitude),Longitude=mean(Longitude))
DJFMP_Coords<-DJFMP_Coords[complete.cases(DJFMP_Coords), ]

# plot the map
map <- create_station_map(DJFMP_Coords) 
map

```

## Subset stations

A lot of stations outside of our study region and many of those are not sampled consistently since the 1970s (based on experience). 

Will subset to just the stations found in my paper:

Mahardja B, Farruggia MJ, Schreier B, Sommer T. 2017. Evidence of a shift in the littoral fish community of the Sacramento-San Joaquin Delta. PLoS One. 12(1):e0170683. https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0170683

See map:

```{r, echo = TRUE, message=FALSE, warning=FALSE}

long_term_stations<- c("DS002S", "GS010E", "LP003E","MK004W", "MR010W", "MS001N","MS001A","OR003W","OR014W","SF014E","SJ001S", "SJ005N", "SJ032S","SJ041N","SJ051E", "SJ056E", "SJ058W", "SJ058E","SR012E","SR012W","SR014W",  "SR017E", "SR024E", "SR043W", "SR049E", "SS011N","TM001N", "WD002W","WD002E","XC001N" )
long_term_stations

DJFMP_Coords_study_stations<-DJFMP_Coords %>% filter(Station %in% long_term_stations)

map <- create_station_map(DJFMP_Coords_study_stations)+ labs(title="All DJFMP stations from Mahardja et al. 2017")
map

```

Check sampling effort for the subset of stations:

```{r, echo = FALSE, message=FALSE, warning=FALSE}

DJFMP_seine_sampling_effort <- DJFMP_Silverside_seine %>% mutate(Year=year(Date),Month=month(Date)) %>%
  group_by(Datetime,Station,Year,Month) %>%
  summarise(Silverside_count=sum(Count)) %>% mutate(sample_size=1) %>% group_by(Station,Year,Month) %>%
  summarise(sample_size=sum(sample_size))%>% filter(Station %in% long_term_stations)

DJFMP_seine_sampling_effort_annual<- DJFMP_seine_sampling_effort %>% group_by(Year,Station) %>% summarise(sample_size=sum(sample_size))

#Plot effort across stations and years for DJFMP beach seine
ggplot(DJFMP_seine_sampling_effort_annual,aes(x=Year,y=as.factor(Station),fill=sample_size))+geom_tile()+
  scale_x_continuous(breaks=c(1970,1980,1990,2000,2010,2015,2020)) + labs(y=NULL,title= "DJFMP station sampling effort by year")


```

Many stations did not start sampling until 1990s. Subset an even smaller number of stations. See map of final stations for **bold**annual data**bold** below. Note that SR012W and SR012E will be used interchangeably because one is meant to replace the other.

```{r, echo = TRUE, message=FALSE, warning=FALSE}

DJFMP_Coords_study_stations_subset_annual <-DJFMP_Coords_study_stations %>% 
  filter(Station %in% c("XC001N","TM001N","SR049E","SR043W","SR024E","SR017E","SR014W","SR012E","SR012W","SJ041N","SJ005N","SJ001S","SF014E","MS001N","MR010W","MK004W","LP003E","GS010E","DS002S"))

DJFMP_Coords_study_stations_subset_annual


map <- create_station_map(DJFMP_Coords_study_stations_subset_annual) + labs(title="DJFMP stations for annual data")
map
```
## Finalize list of stations for annual data

Show sampling effort by month and year for the final list of stations:

```{r, echo = FALSE, message=FALSE, warning=FALSE, fig1, fig.height = 8, fig.width = 9, fig.align = "center"}
DJFMP_seine_sampling_effort_monthly<- DJFMP_seine_sampling_effort %>% filter(Station %in% DJFMP_Coords_study_stations_subset_annual$Station)

ggplot(DJFMP_seine_sampling_effort_monthly,aes(x=Year,y=Month,fill=sample_size))+geom_tile()+facet_wrap(~Station)+
  scale_x_continuous(breaks=c(1970,1980,1990,2000,2010,2015,2020))+ scale_y_continuous(breaks=c(1:12)) + labs(title= "DJFMP station sampling effort by month and year")

```

Lots of months were not sampled in the earlier years (pre-1990s). Sampling timeframe moved from the 1970s to 1990s somewhat. Will use just months from January to May to improve consistency somewhat. WARNING: Note that January to May are when Mississippi Silverside catch is declining (see Mahardja et al. 2016). I didn't include June because two cohorts comingle often in June. 

Mahardja B, Conrad JL, Lusher L, Schreier B. 2016. Abundance Trends, Distribution, and Habitat Associations of the Invasive Mississippi Silverside. San Fr Estuary Watershed Sci. 14(1). doi:10.15447/sfews.2016v14iss1art2. https://escholarship.org/uc/item/55f0s462.

## Final data processing for annual data

* Export stations to data_stations folder
* Join SR012E and SR012W together
* Add length-weight conversion to get biomass
* Use just January-May
* Write csv to annual averages folder


```{r, echo = FALSE, message=FALSE, warning=FALSE}

write.csv((DJFMP_Coords_study_stations_subset_annual %>% mutate(Survey="Delta Juvenile Fish Monitoring Program")),row.names = FALSE,file=file.path("data_stations/Final_stations_fish_DJFMP_annual.csv"))

DJFMP_Silverside_seine_subset<- DJFMP_Silverside_seine
DJFMP_Silverside_seine_subset$Station<-as.factor(DJFMP_Silverside_seine_subset$Station)
levels(DJFMP_Silverside_seine_subset$Station)[levels(DJFMP_Silverside_seine_subset$Station)=="SR012W"] <- "SR012E"

DJFMP_Silverside_seine_subset<- DJFMP_Silverside_seine %>% mutate(Month=month(Date),Year=year(Date),Biomass=ifelse(is.na(Length),0, (0.0097*(Length^2.87))*Count)) %>%
  filter(Station %in% DJFMP_Coords_study_stations_subset_annual$Station & Month %in% c(1:5)) %>%
  group_by(Year, Month, Station, Datetime) %>% summarise(Catch_per_seine=sum(Count),Biomass=sum(Biomass)) %>%
  group_by(Year, Month, Station) %>% summarise(Catch_per_seine=mean(Catch_per_seine),Biomass=mean(Biomass)) %>%
  group_by(Year, Month) %>% summarise(Catch_per_seine=mean(Catch_per_seine),Biomass=mean(Biomass)) %>%
  group_by(Year) %>% summarise(DJFMP_fish_catch_per_seine_Silverside=mean(Catch_per_seine),DJFMP_fish_biomass_Silverside=mean(Biomass))

write.csv(DJFMP_Silverside_seine_subset,row.names=FALSE,file=file.path("annual_averages/fish_DJFMP.csv"))

```

## List of stations for monthly data

Show sampling effort since 1995 for stations
```{r, echo = FALSE, message=FALSE, warning=FALSE}
DJFMP_seine_sampling_effort_monthly<- DJFMP_seine_sampling_effort %>% filter(Station %in% DJFMP_Coords_study_stations$Station, Year >=1995)

ggplot(DJFMP_seine_sampling_effort_monthly,aes(x=Year,y=Month,fill=sample_size))+geom_tile()+facet_wrap(~Station)+
  scale_x_continuous(breaks=c(1995,2000,2005,2010,2015,2020))+ scale_y_continuous(breaks=c(1:12)) + labs(title= "DJFMP station sampling effort by month and year")
```

Combine SJ058E with SJ058W and SR012E with SR012W

```{r, echo = FALSE, message=FALSE, warning=FALSE}
DJFMP_seine_sampling_effort_monthly$Station<-as.factor(DJFMP_seine_sampling_effort_monthly$Station)
levels(DJFMP_seine_sampling_effort_monthly$Station)[levels(DJFMP_seine_sampling_effort_monthly$Station)=="SR012W"] <- "SR012E"
levels(DJFMP_seine_sampling_effort_monthly$Station)[levels(DJFMP_seine_sampling_effort_monthly$Station)=="SJ058E"] <- "SJ058W"

ggplot(DJFMP_seine_sampling_effort_monthly,aes(x=Year,y=Month,fill=sample_size))+geom_tile()+facet_wrap(~Station)+
  scale_x_continuous(breaks=c(1995,2000,2005,2010,2015,2020))+ scale_y_continuous(breaks=c(1:12)) + labs(title= "DJFMP station sampling effort by month and year")
```

All stations look somewhat usable. Export current list of monthly stations to data_stations folder.
```{r, echo = FALSE, message=FALSE, warning=FALSE}

write.csv((DJFMP_Coords_study_stations %>% mutate(Survey="Delta Juvenile Fish Monitoring Program")),row.names = FALSE,file=file.path("stations/Final_stations_fish_DJFMP_monthly.csv"))
```


