---
title: "Salinity model exploration"
author: "Sam Bashevkin"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      knit_root_dir = "../",
      envir = globalenv()
      )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(dplyr)
library(zooper)
library(lubridate)
library(ggplot2)
library(readr)
source("functions/region_assigner.R") #loads function to analyze regional data
```

# Zooplankton

## Data

```{r}
zoop_sal<-zooper::zoopEnvComb%>%
  filter(!is.na(SalSurf) & Source=="EMP" & year(Date)>=1980)%>%
  mutate(salbin=case_when(SalSurf<1 ~ "< 1 PPT",
                          SalSurf>1 & SalSurf <6 ~ "1-6 PPT",
                          SalSurf >6 ~ ">6 PPT"),
         Month=month(Date),
         Year=year(Date),
         salbin=factor(salbin, levels=c("< 1 PPT", "1-6 PPT", ">6 PPT")))%>%
  group_by(Month, Year, salbin)%>%
  summarise(N_stations=n_distinct(Station), .groups="drop")

zoop_geo<-zooper::zoopEnvComb%>%
  filter(!is.na(Latitude) & Source=="EMP" & year(Date)>=1980)%>%
  region_assigner(analysis = "annual", plot = FALSE)%>%
  mutate(Month=month(Date),
         Year=year(Date),
         Region=case_match(Region, "West" ~ "Suisun", "North" ~ "Sacramento", "South" ~ "San Joaquin"),
         Region=factor(Region, levels=c("Suisun", "Sacramento", "San Joaquin")))%>%
  group_by(Month, Year, Region)%>%
  summarise(N_stations=n_distinct(Station), .groups="drop")
```

## Sampling effort plots all dates {.tabset .tabset-pills}


### Salinity {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(zoop_sal, aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~salbin)+
  theme_bw()+
  theme(legend.position = "bottom")
```

### Geography {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(zoop_geo, aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~Region)+
  theme_bw()+
  theme(legend.position = "bottom")
```

## Sampling effort plots 1980 - present {.tabset .tabset-pills}


### Salinity {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(filter(zoop_sal, Year>=1980), aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  geom_text(aes(label=N_stations, color=N_stations<5), size=2, show.legend = F)+
  scale_color_manual(values=c("black", "white"))+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~salbin)+
  theme_bw()+
  theme(legend.position = "bottom")
```

### Geography {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(filter(zoop_geo, Year>=1980), aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  geom_text(aes(label=N_stations, color=N_stations<5), size=2, show.legend = F)+
  scale_color_manual(values=c("black", "white"))+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~Region)+
  theme_bw()+
  theme(legend.position = "bottom")
```

# Benthic Amphipods

## Data

```{r}
df_amph <- read_csv('data/monthly_averages/amphipod_monthly_regions_salinity.csv', col_types = c(Month = 'd')) %>%
  rename(Orig_Region = Region)

df_amph_sal <- df_amph %>%
  filter(!is.na(Salinity) & Year >= 1980)%>%
  mutate(salbin=case_when(Salinity<1 ~ "< 1 PPT",
                          Salinity>1 & Salinity <6 ~ "1-6 PPT",
                          Salinity >6 ~ ">6 PPT"),
         salbin=factor(salbin, levels=c("< 1 PPT", "1-6 PPT", ">6 PPT")))%>%
  group_by(Month, Year, salbin)%>%
  summarise(N_stations=n_distinct(StationCode), .groups="drop")

df_amph_geo <- df_amph %>%
  filter(Year >= 1980)%>%
  region_assigner(analysis = "annual", plot = FALSE)%>%
  mutate(Region=case_when(Orig_Region == "West" ~ "Suisun", Orig_Region == "North" ~ "Sacramento", Orig_Region == "South" ~ "San Joaquin"),
         Region=factor(Region, levels=c("Suisun", "Sacramento", "San Joaquin"))) %>%
  group_by(Month, Year, Region)%>%
  summarise(N_stations=n_distinct(StationCode), .groups="drop")
```

## Sampling effort plots all dates {.tabset .tabset-pills}


### Salinity {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(df_amph_sal, aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~salbin)+
  theme_bw()+
  theme(legend.position = "bottom")
```

### Geography {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(df_amph_geo, aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~Region)+
  theme_bw()+
  theme(legend.position = "bottom")
```

## Sampling effort plots 1980 - present {.tabset .tabset-pills}


### Salinity {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(filter(df_amph_sal, Year>=1980), aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  geom_text(aes(label=N_stations, color=N_stations<5), size=2, show.legend = F)+
  scale_color_manual(values=c("black", "white"))+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~salbin)+
  theme_bw()+
  theme(legend.position = "bottom")
```

### Geography {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(filter(df_amph_geo, Year>=1980), aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  geom_text(aes(label=N_stations, color=N_stations<5), size=2, show.legend = F)+
  scale_color_manual(values=c("black", "white"))+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~Region)+
  theme_bw()+
  theme(legend.position = "bottom")
```


# Benthic Clams

## Data

```{r}
df_clams <- read_csv('data/monthly_averages/clams_monthly_regions_salinity.csv', col_types = c(Month = 'd')) %>%
  rename(Orig_Region = Region)

df_clams_sal <- df_clams %>%
  filter(!is.na(Salinity) & Year >= 1980)%>%
  mutate(salbin=case_when(Salinity<1 ~ "< 1 PPT",
                          Salinity>1 & Salinity <6 ~ "1-6 PPT",
                          Salinity >6 ~ ">6 PPT"),
         salbin=factor(salbin, levels=c("< 1 PPT", "1-6 PPT", ">6 PPT")))%>%
  group_by(Month, Year, salbin)%>%
  summarise(N_stations=n_distinct(StationCode), .groups="drop")

df_clams_geo <- df_clams %>%
  filter(Year >= 1980)%>%
  region_assigner(analysis = "annual", plot = FALSE)%>%
  mutate(Region=case_when(Orig_Region == "West" ~ "Suisun", Orig_Region == "North" ~ "Sacramento", Orig_Region == "South" ~ "San Joaquin"),
         Region=factor(Region, levels=c("Suisun", "Sacramento", "San Joaquin"))) %>%
  group_by(Month, Year, Region)%>%
  summarise(N_stations=n_distinct(StationCode), .groups="drop")
```

## Sampling effort plots all dates {.tabset .tabset-pills}


### Salinity {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(df_clams_sal, aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~salbin)+
  theme_bw()+
  theme(legend.position = "bottom")
```

### Geography {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(df_clams_geo, aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~Region)+
  theme_bw()+
  theme(legend.position = "bottom")
```

## Sampling effort plots 1980 - present {.tabset .tabset-pills}


### Salinity {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(filter(df_clams_sal, Year>=1980), aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  geom_text(aes(label=N_stations, color=N_stations<5), size=2, show.legend = F)+
  scale_color_manual(values=c("black", "white"))+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~salbin)+
  theme_bw()+
  theme(legend.position = "bottom")
```

### Geography {.tabset .tabset-pills}
```{r, fig.width=10, fig.height=8}
ggplot(filter(df_clams_geo, Year>=1980), aes(y=Month, x=Year, fill=N_stations))+
  geom_tile()+
  geom_text(aes(label=N_stations, color=N_stations<5), size=2, show.legend = F)+
  scale_color_manual(values=c("black", "white"))+
  scale_y_continuous(breaks=1:12)+
  scale_fill_viridis_c()+
  facet_wrap(~Region)+
  theme_bw()+
  theme(legend.position = "bottom")
```
