---
title: "Data_wrangling_fish"
author: "Denise-Colombano"
date: "10/28/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to evaluate the spatial and temporal overlap of 
the different surveys used in the SEM.

Inventory datasets:
- Fish: FMWT, Bay Study
- Zoop: EMP
- [TBD]


# Load libraries
```{r}
# tidy data
library(tidyverse)

# S.B. packages
library(zooper)
library(LTMRdata)
library(discretewq)

# J.C. packages
library(DT)
library(leaflet)
library(sf)
library(janitor)

```

# Fish surveys

Query the CDFW Fall Midwater Trawl and Bay Study data sets from the LTMRdata package.
```{r}
Surveys <- LTMRdata::fish(sources=c("FMWT", "Baystudy"),
             species=NULL,
             convert_lengths=TRUE,
             remove_unconverted_lengths=FALSE,
             size_cutoff=40,
             zero_fill = TRUE,
             remove_unknown_lengths = TRUE,
             univariate = FALSE,
             quiet = FALSE)
Surveys

Surveys$Month <- lubridate::month(Surveys$Date)
Surveys$Year <- lubridate::year(Surveys$Date)
str(Surveys)
```


Assess sampling density across space and time
```{r}
SamplingDensity <- Surveys %>%
  select(Source, Method, Station, Latitude, Longitude, Year, Month) %>% 
  distinct(Source, Method, Station, Latitude, Longitude, Year, Month) %>% 
  mutate(Sampling=1)

summary(SamplingDensity)

# Bay Study
Baystudy <- SamplingDensity %>% 
  filter(Source=="Bay Study") %>% 
  group_by(Station, Longitude, Latitude, Year) %>% 
  summarize(Sampling=sum(Sampling)/24) # 2 gears x 12 months = 24 sampling points per year
Baystudy

ggplot(Baystudy, aes(Longitude, Latitude))+
  geom_point(aes(color=Station, size=Sampling), pch=21)+
  theme_bw()+
  facet_wrap(Year~.)+
  scale_color_viridis_d()

# FMWT
Fallmwt <- SamplingDensity %>% 
  filter(Source=="FMWT") %>% 
  group_by(Station, Longitude, Latitude, Year) %>% 
  summarize(Sampling=sum(Sampling)) 
Fallmwt

ggplot(Fallmwt, aes(Longitude, Latitude))+
  geom_point(aes(color=Station, size=Sampling), pch=21)+
  theme_bw()+
  facet_wrap(Year~.)+
  scale_color_viridis_d()
```


# Zoop surveys

Query the EMP data from the Zooper package. Borrow code from "zoop.Rmd" by S.B.

Load core stations
```{r, message=FALSE}
download.file("https://portal.edirepository.org/nis/dataviewer?packageid=edi.522.7&entityid=71dd301f30a2bc2e40f5da573dde9f97", destfile = file.path(tempdir(), "zoop_station_lookup.csv"))

zoop_stations<-read_csv(file.path(tempdir(), "zoop_station_lookup.csv"))%>%
  filter(Core%in%c(1, 2))
```

Load zoop data
```{r}
zoop_data<-Zoopsynther(Data_type="Community", Sources="EMP", Time_consistency = TRUE)%>%
  filter(Station%in%zoop_stations$StationNZ)%>%
  mutate(Month=lubridate::month(Date))
```

Remove stations that aren't continuous
```{r}
zoop_data_continuous<-filter(zoop_data, !Station%in%c("NZEZ6", "NZEZ2", "NZD16", "NZD06", "NZ080", "NZ042"))

# Make sure the right number of stations was removed
length(setdiff(unique(zoop_data$Station), unique(zoop_data_continuous$Station)))==6

str(zoop_data_continuous)
```

Assess sampling density across space and time
```{r}
zoop_surveys <- zoop_data_continuous %>% 
  select(Month, Year, Station, Longitude, Latitude) %>% 
  distinct(Month, Year, Station, Longitude, Latitude) %>% 
  mutate(Sampling=1)

zoop_emp <- zoop_surveys %>% 
  group_by(Station, Year, Longitude, Latitude) %>% 
  summarize(Sampling=sum(Sampling))

ggplot(zoop_emp, aes(Longitude, Latitude))+
  geom_point(aes(color=Station, size=Sampling), pch=21)+
  theme_bw()+
  facet_wrap(Year~.)+
  scale_color_viridis_d()
```

# Fish + Zoop
Layer them on top of each other
```{r}
# map for each year
ggplot()+
  geom_point(data=Fallmwt, aes(Longitude, Latitude), pch=21, color="gray50")+
  geom_point(data=Baystudy, aes(Longitude, Latitude), pch=21, color="orange")+
  geom_point(data=zoop_emp, aes(Longitude, Latitude), pch=4, color="slateblue")+  
  theme_bw()+
  facet_wrap(Year~.)+
  labs(subtitle="Spatial overlap - annual")

# map for all years
ggplot()+
  geom_point(data=Fallmwt, aes(Longitude, Latitude), pch=21, color="gray50", size=3)+
  geom_point(data=Baystudy, aes(Longitude, Latitude), pch=21, color="orange", size=3)+
  geom_point(data=zoop_emp, aes(Longitude, Latitude), pch=3, color="slateblue", size=3)+  
  theme_bw()+
  labs(subtitle="Spatial overlap - overall")
```


# Water quality

Same stations as EMP zooplankton that are already plotted.
Data lives here.
```{r}
# require(discretewq)
# wq_stations <- wq(Sources = "EMP")
# wq_stations <- pivot_longer(wq_stations, 
#                             cols = c(Temperature, Chlorophyll:TKN, Salinity), 
#                             names_to = 'Analyte', values_to = 'Value')
```


# Benthic 

Code borrowed from the benthic.Rmd by JC
```{r, message=FALSE}
ben <- read_csv("data_raw/DWR Benthic raw count data 1975-2020 2021_10_01.csv")  %>% 
  mutate(StationClean =  gsub("-[A-Z]","",StationCode)) %>% 
  mutate(mon = lubridate::ym(paste0(Year, Month, sep = "-"))) %>% 
  group_by(StationClean) %>% 
  mutate(Latitude = mean(Latitude, na.rm = T), 
         Longitude = mean(Longitude, na.rm = T)) %>% 
  ungroup()
```

```{r}
sites_mon <- ben  %>% 
  distinct(StationClean, Latitude, Longitude, mon) 

sites <- ben %>% 
  group_by(StationClean, Latitude, Longitude) %>% 
  summarise(n_months = length(unique(mon)), .groups  =  "drop") %>% 
  st_as_sf(coords = c('Longitude','Latitude'), crs = 4326,  remove = FALSE)
```

## Sites that were sampled each month


```{r, echo=F}
ggplot(sites_mon, aes(x = mon, y = StationClean), size = 3) +
  geom_point() +
  theme_classic()
```


## Table of site information

```{r, echo = FALSE}
datatable(as.data.frame(sites))
```

## Core Stations

```{r}
core <- c("P8",
          "MD10",
          "D8",
          "D7",
          "D6",
          "D41",
          "D4",
          "D28A",
          "D26",
          "C3",
          "C10")

core_data <-  sites_mon %>% 
  filter(StationClean %in% core)

sites_core <- sites %>% 
  filter(StationClean %in% core)
```


```{r, echo=F}
ggplot(core_data, aes(x = mon, y = StationClean), size = 3) +
  geom_point() +
  theme_classic()
```



# Plot altogether
```{r, message = FALSE, echo = FALSE}

leaflet(sites) %>% 
  addTiles() %>% 
  addCircleMarkers(data = Fallmwt,
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = 2, 
                         fillColor = "gray",
                         fillOpacity = 0.5,
                         weight = 0.1,
                         color = "black",
                         label = ~Station) %>% 
  addCircleMarkers(data = Baystudy,
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = 2,
                         fillColor = "yellow",
                         fillOpacity = 0.5,
                         weight = 0.1,
                         color = "black",
                         label = ~Station) %>% 
  addCircleMarkers(data = zoop_emp,
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = 2,                         fillColor = "blue",
                         fillOpacity = 1,
                         weight = 0.1,
                         color = "black",
                         label = ~Station) %>% 
  addCircleMarkers(data = sites_core,
                         lat = ~Latitude,
                         lng = ~Longitude,
                         radius = 3, 
                         fillColor = "red",
                         fillOpacity = 1,
                         weight = 0.1,
                         color = "black",
                         label = ~StationClean)
```

