---
title: "zoop heatmap"
output: html_document
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      knit_root_dir = "../",
      envir = globalenv()
    )
    })
---

TODO:
1) Need to get energy density to work with Mysids. 
Need to remove Mysids from the dry weights, and add it with energy density but just as an overall Mysid category

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r, message=FALSE}
require(dplyr)
require(zooper)
require(lubridate)
require(readr)
require(tidyr)
require(ggplot2)
require(readxl)
require(stringr)
require(sf)
source("functions/region_assigner.R")
```

Load core stations
```{r, message=FALSE}
download.file("https://portal.edirepository.org/nis/dataviewer?packageid=edi.522.7&entityid=71dd301f30a2bc2e40f5da573dde9f97", destfile = file.path(tempdir(), "zoop_station_lookup.csv"))

zoop_stations<-read_csv(file.path(tempdir(), "zoop_station_lookup.csv"))%>%
  filter(Core%in%c(1, 2))
```

Load zoop data
```{r}
zoop_data<-Zoopsynther(Data_type="Community", Sources="EMP", Time_consistency = TRUE)%>%
  mutate(Month=month(Date))
```


Read in zoop mass conversions and groupings
Also Read in the energy density data and join to the mass grouping
```{r}
zoop_energy_data<-read.csv("data/zoop_energy_data/taxa_energy_density_dry.csv")

zoop_mass_group<-read_excel("data/data_in/Zoop Categories for SEM.xlsx", na = "NA")%>%
  mutate(Taxlifestage=paste(Taxname, Lifestage))%>%
  rename(Group=`SEM Category`)%>%
  mutate(Group=str_replace_all(Group, " ", "_"))%>%
  filter(!is.na(Carbon_mass_micrograms)) %>% 
  left_join(zoop_energy_data, by = "Taxname")%>%
  mutate(Taxlifestage=if_else(Group=="Mysids", "Mysida", Taxlifestage))%>%
  select(Group, Taxlifestage, Carbon_mass_micrograms, energy_density_j_per_g_dry_mass)

# List all the zoop groups
unique(zoop_mass_group$Group)
```

Load Mysid biomass data
```{r}

mysid_energy = zoop_energy_data %>% 
  filter(Taxonomic.Category == "Mysids") %>% 
  summarise(energy = mean(energy_density_j_per_g_dry_mass)) %>% 
  unlist()
  

zoop_mysid<-read_excel("data/data_in/1972-2020MysidBPUEMatrix.xlsx",
                       sheet="Mysid_BPUE_matrix_1972-2020", na = "NA",
                       col_types = c(rep("numeric", 4), "date", "text", "text", rep("text", 7), rep("numeric", 8)))%>%
  select(Date=SampleDate, Station=StationNZ, `Acanthomysis aspera`:Unidentified)%>%
  mutate(BPUE_mysid=rowSums(select(., -Date, -Station), na.rm=T))%>%
  mutate(BPUE_mysid=BPUE_mysid*1000, # Convert to ug
         Group="Mysids")%>%
  mutate(SampleID=paste("EMP", Station, Date))%>%
  select(SampleID, Group, BPUE_mysid)
```


Start processing the zoop data
```{r}
zoop_data_mass<-zoop_data%>%
  mutate(Taxlifestage=str_remove(Taxlifestage, fixed("_UnID")))%>%
  filter(
    !(SizeClass=="Meso" & #eliminating species which are counted in meso and micro and retained better in the micro net from the meso calcs
        
        Taxlifestage%in%c("Asplanchna Adult", "Copepoda Larva","Cyclopoida Juvenile", "Eurytemora affinis Larva", "Harpacticoida Undifferentiated",
                          "Keratella Adult", "Limnoithona Adult", "Limnoithona Juvenile", "Limnoithona sinenesis Adult", "Limnoithona tetraspina
                                    Adult", "Oithona Adult", "Oithona Juvenile", "Oithona davisae Adult", "Polyarthra Adult","Pseudodiaptomus Larva", 
                          "Rotifera Adult", "Sinocalanus doerrii Larva", "Synchaeta Adult", "Synchaeta bicornis Adult", "Trichocerca Adult")) &
      
      !(SizeClass=="Micro" & #removing categories better retained in meso net from micro net matrix
          
          Taxlifestage%in%c("Cirripedia Larva", "Cyclopoida Adult", "Oithona similis")))%>%
  mutate(Taxlifestage=if_else(Order=="Mysida", "Mysida", Taxlifestage), # Need to summarise Mysid CPUE first before joining to Mysid BPUE
         Taxname=if_else(Order=="Mysida", "Mysida", Taxname))%>%
  group_by(across(-c(Species, Genus, Family, CPUE)))%>%
  summarise(CPUE=sum(CPUE), .groups="drop")%>%
  left_join(zoop_mass_group, by=c("Taxlifestage"))%>%
  mutate(BPUE=CPUE*Carbon_mass_micrograms,
         Group=if_else(Order=="Mysida", "Mysids", Group))%>%
  left_join(zoop_mysid, 
            by=c("Group", "SampleID"))%>%
  filter(!is.na(Group))%>% # This removes anyone without an assigned group
  mutate(BPUE=if_else(is.na(BPUE), BPUE_mysid, BPUE),
         energy_density_j_per_g_dry_mass = if_else(Group == "Mysids",
                                                   mysid_energy,
                                                   energy_density_j_per_g_dry_mass),
         JPUE=(BPUE/1000)*energy_density_j_per_g_dry_mass)
```



Summarize zoop effort by month and year
```{r}
zoop_data_sum<-zoop_data_mass%>%
  select(Month, Year, SampleID)%>%
  distinct()%>%
  group_by(Month, Year)%>%
  summarise(N=n(), .groups="drop")
```

Plot heat map
```{r}
ggplot(zoop_data_sum, aes(y=Month, x=Year, fill=N))+
  geom_tile()+
  scale_fill_viridis_c()+
  scale_y_continuous(breaks=1:12)+
  coord_cartesian(expand=FALSE)+
  theme_bw()
```

Summarize zoop effort by date and station
```{r}
zoop_data_station<-zoop_data_mass%>%
  select(Month, Year, Station)%>%
  distinct()%>%
  mutate(type="present")%>%
  complete(Month, Year, Station, fill=list(type="missing"))%>%
  mutate(Date=dmy(paste("1", Month, Year)))
```

Plot zoop effort by date and station for annual metrics
```{r}
ggplot(zoop_data_station%>%filter(Station%in%zoop_stations$StationNZ), aes(x=Date, y=Station, color=type))+
  geom_point()+
  scale_color_manual(values=c("firebrick3", "black"))+
  theme_bw()
```

Find zoop stations that can be used for monthly metrics
```{r}
zoop_stations_month<-zoop_data_mass%>%
  filter(Year>=1995)%>%
  group_by(Station, Month)%>%
  summarise(N_years=n_distinct(Year), .groups="drop")%>%
  filter(N_years>=22)%>%
  group_by(Station)%>%
  summarise(N_months=n_distinct(Month))%>%
  filter(N_months==12)
```

Plot zoop effort by date and station for monthly metrics
```{r}
ggplot(zoop_data_station%>%filter(Station%in%zoop_stations_month$Station), aes(x=Date, y=Station, color=type))+
  geom_point()+
  scale_color_manual(values=c("firebrick3", "black"))+
  theme_bw()
```

Create annual data set with final set of annual stations
```{r}
zoop_data_annual<-zoop_data_mass%>%
  filter(Station%in%zoop_stations$StationNZ)%>%
  filter(!Station%in%c("NZEZ6", "NZEZ2", "NZD16", "NZD06", "NZ080", "NZ042"))%>% # Renove stations that aren't continuous in time
  filter(Month%in%3:11)

# Save list of annual stations
zoop_data_annual%>%
  select(Station, Latitude, Longitude)%>%
  mutate(Survey="EMP zooplantkon")%>%
  distinct()%>%
  write_csv("data/stations/stations_zoop_annual.csv")
```

Plot the final set of annual stations
```{r}
ggplot()+
  geom_sf(data=deltamapr::WW_Delta%>%st_transform(crs=4326))+
  geom_point(data=zoop_data_annual%>%select(Station, Latitude, Longitude)%>%distinct(), 
             aes(x=Longitude, y=Latitude),
             color="red")
```

Create monthly data set with final set of monthly stations
```{r}
zoop_data_month<-zoop_data_mass%>%
  filter(Station%in%zoop_stations_month$Station)

# Save list of annual stations
zoop_data_month%>%
  select(Station, Latitude, Longitude)%>%
  mutate(Survey="EMP zooplantkon")%>%
  distinct()%>%
  write_csv("data/stations/stations_zoop_month.csv")
```

Plot the final set of monthly stations
```{r}
ggplot()+
  geom_sf(data=deltamapr::WW_Delta%>%st_transform(crs=4326))+
  geom_point(data=zoop_data_month%>%select(Station, Latitude, Longitude)%>%distinct(), 
             aes(x=Longitude, y=Latitude),
             color="red")
```

# Export zoop data for analysis

Annual data
```{r}
zoop_data_annual_final<-zoop_data_annual%>%
  region_assigner(analysis="annual")

zoop_data_annual_final_noregions<-zoop_data_annual_final%>%
  group_by(Year, Group)%>%
  summarise(across(c(BPUE, CPUE, JPUE), ~sum(.x, na.rm=T)), .groups="drop")%>%
  pivot_wider(names_from = Group, values_from = c(BPUE, CPUE, JPUE),
              names_glue = "{Group}_{.value}")

str(zoop_data_annual_final_noregions)
write_csv(zoop_data_annual_final_noregions, "data/annual_averages/zoop_annual_noregions.csv")

zoop_data_annual_final_regions<-zoop_data_annual_final%>%
  group_by(Year, Group, Region)%>%
  summarise(across(c(BPUE, CPUE, JPUE), ~sum(.x, na.rm=T)), .groups="drop")%>%
  pivot_wider(names_from = Group, values_from = c(BPUE, CPUE, JPUE),
              names_glue = "{Group}_{.value}")

str(zoop_data_annual_final_regions)
write_csv(zoop_data_annual_final_regions, "data/annual_averages/zoop_annual_regions.csv")
```

Monthly data
```{r}
zoop_data_month_final<-zoop_data_month%>%
  region_assigner(analysis="monthly")%>%
  group_by(Year, Month, Group, Region)%>%
  summarise(across(c(BPUE, CPUE, JPUE), ~sum(.x, na.rm=T)), .groups="drop")%>%
  pivot_wider(names_from = Group, values_from = c(BPUE, CPUE, JPUE),
              names_glue = "{Group}_{.value}")

str(zoop_data_month_final)

write_csv(zoop_data_month_final, "data/monthly_averages/zoop_month.csv")
```


