---
title: "Zoo Energy Densities"
author: "Peter Dudley"
date: "10/29/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      knit_root_dir = "../",
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Get Zoo energy density
This script is to analyze the energy density of the categories of zooplankton used in out SEM food web model. 

# Libraries and sourced functions used
```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)

source("./data/zoop_energy_functions.R")
```

# Load the data
All the files are generated with a excel macro so their file names all start the same way. I use map to read them all into a list
```{r load_data, }
data = load_zoop_energy_density("./data/zoop_energy_data")
```

# Process Data
If there are groups that are larger then our categories, make entries for each sub
category containing that data the rows for each one
```{r}
summary_ratio = summarize_zoop(data$ratio,
                             "dry_to_wet_ratio")

# convert wet to dry if data available
converted_dry = convert_zoop_energy_density(ratio_df = summary_ratio,
                                            energy_df = data$wet,
                                            convert_to = "dry") %>% 
  bind_rows(data$dry)

converted_wet = convert_zoop_energy_density(ratio_df = summary_ratio,
                                            energy_df = data$dry,
                                            convert_to = "wet") %>% 
  bind_rows(data$wet)

```

Convert all to wet weight and print a table
```{r}

wet_summary = summarize_zoop(converted_wet,
                             "energy_density_j_per_g_wet_mass") 
print(wet_summary)

write.csv(wet_summary, file ="./data/zoop_energy_data/zoop_energy_density_wet.csv")

```
Convert all to day weight and print a table
```{r}

dry_summary = summarize_zoop(converted_dry,
                             "energy_density_j_per_g_dry_mass") 
print(dry_summary)

write.csv(dry_summary, file ="./data/zoop_energy_data/zoop_energy_density_dry.csv")

```

# Plots
Print a box plot of the data
```{r fig1, fig.height = 8, fig.width = 5}
density_plot = ggplot(data = converted_wet,
                      aes(x = fct_reorder(group, energy_density_j_per_g_wet_mass),
                          y = energy_density_j_per_g_wet_mass)) + 
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "Group",
       y = "Energy Density (j/g wet mass)" ) +
  geom_boxplot() +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=2, binwidth = 30, alpha = 0.2)+
  stat_summary(fun=mean, geom="point", color = "red", shape=23, size=4, stroke = 2)
print(density_plot)
```

