---
title: "Zoo Energy Densities"
author: "Peter Dudley"
date: "10/29/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Get Zoo energy density
This script is to analyze the energy density of the categories of zooplankton used in out SEM food web model. 

# Libraries and sourced functions used
```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)

source("./breakout_zoop_energy_groups.R")
```

# Load the data
All the files are generated with a excel macro so their file names all start the same way. I use map to read them all into a list
```{r load_data}
file_name_start = "./zoop_energy_data/zoo_energy_"
catagories = na.omit(read.csv(paste0(file_name_start, "catagories.csv"))) %>% 
  mutate_if(is.character, str_trim)
file_names = as.list(c( "dry", "wet", "digestibility", "ratio"))
data = file_names %>% 
  map(~na.omit(read.csv(paste0(file_name_start, .x, ".csv"))) %>% 
        mutate_if(is.character, str_trim)) %>% 
  set_names(file_names)
```

# Process Data
If there are groups that are larger then our categories, make entries for each sub
category containing that data the rows for each one
```{r}
higher_groups = catagories %>% 
  na.omit()

# break out the catagories
data_brokenout = data %>% 
  map(~break_out_groups(.x, higher_groups, "group"))

data_ratio = data_brokenout$ratio %>% 
  group_by(group) %>% 
  summarize(dry_to_wet_ratio = mean(dry_to_wet_ratio, na.rm = TRUE))

# convert wet to dry if data available
converted_dry = data_ratio %>% 
  na.omit() %>% 
  right_join(data_brokenout$wet, by = "group") %>% 
  mutate(energy_density_j_per_g_dry_mass =
           energy_density_j_per_g_wet_mass/dry_to_wet_ratio) %>% 
  select(-energy_density_j_per_g_wet_mass, -dry_to_wet_ratio) %>%
  bind_rows(data_brokenout$dry)

converted_wet = data_ratio %>% 
  na.omit() %>% 
  right_join(data_brokenout$dry, by = "group") %>% 
  mutate(energy_density_j_per_g_wet_mass =
           energy_density_j_per_g_dry_mass*dry_to_wet_ratio) %>% 
  select(-energy_density_j_per_g_dry_mass, -dry_to_wet_ratio) %>%
  bind_rows(data_brokenout$wet)

```

Convert all to wet weight and print a table
```{r}
dry_summary = converted_wet %>% 
  group_by(group) %>% 
  summarise(mean_energy_density = mean(energy_density_j_per_g_wet_mass),
            energy_density_sd = sd(energy_density_j_per_g_wet_mass))
print(dry_summary)


```
# Plots
Print a box plot of the data
```{r fig1, fig.height = 8, fig.width = 5}
density_plot = ggplot(data = converted_wet,
                      aes(x = fct_reorder(group, energy_density_j_per_g_wet_mass),
                          y = energy_density_j_per_g_wet_mass)) + 
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "Group",
       y = "Energy Density (j/g wet mass)" ) +
  geom_boxplot() +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=2, binwidth = 30, alpha = 0.2)+
  stat_summary(fun=mean, geom="point", color = "red", shape=23, size=4, stroke = 2)
print(density_plot)
```

