---
title: "Preliminary SEM Analyses 2"
author: "Delta Food Web Team"
date: "10/25/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      knit_root_dir = '..',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(dplyr)
library(psych) #for pairs.panels, but could use other packages, e.g. GGalley
library(lavaan)
library(semPlot)
```

### Import data

```{r dataprep}
combined=read.csv("data/annual_averages/annual_data_compiled.csv")
cnames=read.csv("analysis/column_names.csv", stringsAsFactors = F)
dsub=filter(combined, Year>=1975)
focaldata=dsub[,cnames$Datacolumn]
fvars=cnames$Shortname
colnames(focaldata)=fvars
```

### Derived variables

```{r}
focaldata = focaldata %>% 
  mutate(hzoop=hcope+clad) #biomass, herbivorous zooplankton
fvars=c(fvars,"hzoop")
```

### Data prep

Log transform, scale

```{r histo}
#log transform
logvars=fvars[c(cnames$Log=="yes",T)]
logtrans=function(x) {
  x2=x[which(!is.na(x))]
  if(any(x2==0)) {log(x+min(x2[which(x2>0)],na.rm=T))}
  else {log(x)}
}
focaldatalog = focaldata %>% 
  mutate_at(logvars,logtrans)

#scale data
fd=focaldatalog
fd=fd %>% 
  mutate_at(2:length(fvars),list("1"=lag)) %>% #lag 1
  mutate_at(2:length(fvars),list("fd"=function(x) c(NA,diff(x)))) %>% #first difference
  mutate_at(2:length(fvars),list("dtr"=function(x) { #detrend
    x2=x
    x2[x2==0]=NA
    res=residuals(lm(x2~fd$year))
    out=x
    out[which(!is.na(x2))]=res
    return(out)
  })) %>%
  mutate_at(-1,scale)
```

### Time series plots

Original units

```{r timeseries1, fig.width=8, fig.height=11, echo=FALSE}
#original units
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=focaldata[,1],y=focaldata[,i],main=fvars[i], type="o",cex=0.5)
}
```

Log scaled

```{r timeseries2, fig.width=8, fig.height=11, echo=FALSE}
#log scaled
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i],main=fvars[i], type="o",cex=0.5)
}
```

First difference

```{r timeseries3, fig.width=8, fig.height=11, echo=FALSE}
#first diff
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i+2*(length(fvars)-1)],main=fvars[i], type="o",cex=0.5)
}
```

Detrended

```{r timeseries4, fig.width=8, fig.height=11, echo=FALSE}
#detrended
par(mfrow=c(7,5),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i+3*(length(fvars)-1)],main=fvars[i], type="o",cex=0.5)
}
```

Bivariate plots

```{r bivariate, fig.width=8, fig.height=8, echo=FALSE}
pairs.panels(select(fd, year, flow, temp, secchi, chla, hzoop, pcope, mysid, smelt),lm = T)
```

### SEM model

Simplified model with just one category for herbiv zoop.

#### Smelt

```{r}
model1='chla~year+flow+secchi
        hzoop~year+flow+temp+chla+pcope+secchi
        pcope~year+flow
        mysid~year+flow+hzoop+pcope
        smelt~year+flow+secchi+hzoop+pcope+mysid'
modfit1=sem(model1, data=fd)
#summary(modfit1, standardized=T, rsq=T)
semPaths(modfit1, "std", edge.label.cex = 1, residuals = F)

#delete some of ns paths
model1b='chla~year+flow+secchi
        hzoop~year+flow+temp+chla+pcope+secchi
        pcope~year
        mysid~year+hzoop+pcope
        smelt~year+secchi+hzoop+pcope+mysid'
modfit1b=sem(model1b, data=fd)
#summary(modfit1b, standardized=T, rsq=T)
semPaths(modfit1b, "std", edge.label.cex = 1, residuals = F)

anova(modfit1, modfit1b)

# #adding smelt last year makes year effects ns...
# model1c='chla~year+flow+secchi
#         hzoop~year+flow+temp+chla+pcope+secchi
#         pcope~year
#         mysid~year+hzoop+pcope
#         smelt~year+secchi+hzoop+pcope+mysid+smelt_1'
# modfit1c=sem(model1c, data=fd)
# summary(modfit1c, standardized=T, rsq=T)
# semPaths(modfit1c, "std", edge.label.cex = 1, residuals = F)

# #year vs secchi effect on hzoop
# model1d='chla~year+flow+secchi
#         hzoop~flow+temp+chla+pcope+secchi
#         pcope~year
#         mysid~year+hzoop+pcope
#         smelt~year+secchi+hzoop+pcope+mysid'
# modfit1d=sem(model1d, data=fd)
# summary(modfit1d, standardized=T, rsq=T)
# semPaths(modfit1d, "std", edge.label.cex = 1, residuals = F)
# 
# anova(modfit1d, modfit1b)

#using hcope instead of hzoop
model1e='chla~year+flow+secchi
        hcope~year+flow+temp+chla+pcope+secchi
        pcope~year
        mysid~year+hcope+pcope
        smelt~year+secchi+hcope+pcope+mysid'
modfit1e=sem(model1e, data=fd)
#summary(modfit1e, standardized=T, rsq=T)
semPaths(modfit1e, "std", edge.label.cex = 1, residuals = F)
```

#### Fish aggregates

```{r}
#using estfish instead of smelt
model1f='chla~year+flow+secchi
        hcope~year+flow+temp+chla+pcope+secchi
        pcope~year
        mysid~year+hcope+pcope
        estfish~year+secchi+hcope+pcope+mysid'
modfit1f=sem(model1f, data=fd)
#summary(modfit1f, standardized=T, rsq=T)
semPaths(modfit1f, "std", edge.label.cex = 1, residuals = F)

#using marfish instead of smelt
model1g='chla~year+flow+secchi
        hcope~year+flow+temp+chla+pcope+secchi
        pcope~year
        mysid~year+hcope+pcope
        marfish~year+secchi+hcope+pcope+mysid'
modfit1g=sem(model1g, data=fd)
#summary(modfit1g, standardized=T, rsq=T)
semPaths(modfit1g, "std", edge.label.cex = 1, residuals = F)
```

Detrended (anomalies) without year effect

```{r}
#smelt
model2a='chla_dtr~flow_dtr+temp_dtr+secchi_dtr
         hzoop_dtr~flow_dtr+temp_dtr+secchi_dtr+chla_dtr+pcope_dtr
         pcope_dtr~flow_dtr+temp_dtr+secchi_dtr
         mysid_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr
         smelt_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr+mysid_dtr'
modfit2a=sem(model2a, data=fd)
#summary(modfit2a, standardized=T, rsq=T)
semPaths(modfit2a, "std", edge.label.cex = 1, residuals = F)

#estuarine fishes
model2b='chla_dtr~flow_dtr+temp_dtr+secchi_dtr
         hzoop_dtr~flow_dtr+temp_dtr+secchi_dtr+chla_dtr+pcope_dtr
         pcope_dtr~flow_dtr+temp_dtr+secchi_dtr
         mysid_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr
         estfish_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr+mysid_dtr'
modfit2b=sem(model2b, data=fd)
#summary(modfit2b, standardized=T, rsq=T)
semPaths(modfit2b, "std", edge.label.cex = 1, residuals = F)
```

Estuarine fishes with latent variable, detrended 

```{r}
#estuarine fishes with latent variable
model2c='chla_dtr~flow_dtr+temp_dtr+secchi_dtr
         hzoop_dtr~flow_dtr+temp_dtr+secchi_dtr+chla_dtr+pcope_dtr
         pcope_dtr~flow_dtr+temp_dtr+secchi_dtr
         mysid_dtr~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr
         estfish_lat~flow_dtr+temp_dtr+secchi_dtr+hzoop_dtr+pcope_dtr+mysid_dtr
         estfish_lat=~estfish_dtr+estfish_bsmt_dtr' #+estfish_bsot_dtr
modfit2c=sem(model2c, data=fd)
#summary(modfit2c, standardized=T, rsq=T)
semPaths(modfit2c, "std", edge.label.cex = 1, residuals = F, layout = "tree2")
```

Estuarine fishes with latent variable, not detrended

```{r}
#estuarine fishes with latent variable (not detrended with year)
model2d='chla~year+flow+temp+secchi
         hzoop~year+flow+temp+secchi+chla+pcope
         pcope~year+flow+temp+secchi
         mysid~year+flow+temp+secchi+hzoop+pcope
         estfish_lat~year+flow+temp+secchi+hzoop+pcope+mysid
         estfish_lat=~estfish+estfish_bsmt' #+estfish_bsot
modfit2d=sem(model2d, data=fd)
#summary(modfit2d, standardized=T, rsq=T)
semPaths(modfit2d, "std", edge.label.cex = 1, residuals = F, layout = "tree2")
```

