---
title: "Monthly model - Regions"
author: "Tanya Rogers"
date: "1/6/2022"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      knit_root_dir = '..',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(dplyr)
#library(psych) #for pairs.panels, but could use other packages, e.g. GGalley
library(lavaan)
library(semPlot)
library(DiagrammeR)
library(ggplot2)
library(tidyr)
```

### Import data

```{r import}
combined=read.csv("data/monthly_averages/monthly_data_compiled_regions.csv",stringsAsFactors = F)
cnames=read.csv("analysis/column_names_region_monthly.csv", stringsAsFactors = F)
dsub=filter(combined, Year>=1995) %>% arrange(Region,Year,Month)
focaldata=dsub[,cnames$Datacolumn]
fvars=cnames$Shortname
colnames(focaldata)=fvars
regions=unique(focaldata$region)
regionorder=c("Far West","West","North","South")
focaldata=focaldata%>% 
  mutate(decyear=year+(month-1)/12,
         sinmon=sin(2*pi/12*month),
         cosmon=cos(2*pi/12*month))

focaldata = focaldata %>% 
  mutate(tzoop=hcope+clad+mysid+pcope,
         tzoop_c=hcope_c+clad_c+mysid_c+pcope_c,
         tzoop_e=hcope_e+clad_e+mysid_e+pcope_e,
         hzoop=hcope+clad,
         hzoop_c=hcope_c+clad_c,
         hzoop_e=hcope_e+clad_e,
         pzoop=mysid+pcope,
         pzoop_c=mysid_c+pcope_c,
         pzoop_e=mysid_e+pcope_e) 
fvars=c(fvars,"tzoop","tzoop_c","tzoop_e",
        "hzoop","hzoop_c","hzoop_e",
        "pzoop","pzoop_c","pzoop_e")
cnames=rbind(cnames,data.frame(Longname=NA,Shortname=c("tzoop","tzoop_c","tzoop_e",
                                                       "hzoop","hzoop_c","hzoop_e",
                                                       "pzoop","pzoop_c","pzoop_e"),
                               Diagramname=c("Total Zooplankton\nBiomass",
                                             "Total Zooplankton\nAbundance",
                                             "Total Zooplankton\nEnergy",
                                             "Herbivorous Zooplankton\nBiomass",
                                             "Herbivorous Zooplankton\nAbundance",
                                             "Herbivorous Zooplankton\nEnergy",
                                             "Predatory Zooplankton\nBiomass",
                                             "Predatory Zooplankton\nAbundance",
                                             "Predatory Zooplankton\nEnergy"),
                               Datacolumn=NA,Log="yes"))

#focal variables
varnames=c("temp","flow","nitrate","ammonia","dophos","chla","hcope","clad","amphi","pcope","mysid","potam","corbic","sside","estfish_bsot","estfish_bsmt","tzoop","tzoop_c","tzoop_e","hzoop","pzoop")

#labels for lagged vars
cnameslag=cnames
cnameslag$Shortname=paste0(cnameslag$Shortname,"_1")
cnameslag$Diagramname=paste(cnameslag$Diagramname,"(t-1)")
cnameslag=rbind(cnames,cnameslag)

source("analysis/myLavaanPlot.r")
```

Notes:  
Determine which vars are appropriate for which regions and at what lags.  
Write out path diagram for each region.  
Model with and without fish, probably, because of missing data.  

### Data prep

Log transform, scale.  
Within and across regions.  
Create set with regional monthly means removed.  

```{r prep}
#log transform
logvars=fvars[cnames$Log=="yes"]
logtrans=function(x) {
  x2=x[which(!is.na(x))]
  if(any(x2==0)) {log(x+min(x2[which(x2>0)],na.rm=T))}
  else {log(x)}
}
focaldatalog = focaldata %>% 
  mutate_at(logvars,logtrans)

#scale data
fdr0=focaldatalog
tvars=fvars[-(1:3)]

#scaled within regions
fdr=fdr0 %>% 
  group_by(region) %>% 
  #scale
  mutate_at(tvars,scale) %>% 
  #lag
  mutate_at(tvars,list("1"=lag,"2"=function(x) {lag(x,2)})) %>% 
  ungroup() %>% 
  as.data.frame()

#scaled within regions, remove monthly means
fdr_ds=fdr %>% 
  group_by(region,month) %>%
  mutate_at(tvars,list("mm"=function(x) {mean(x,na.rm = T)})) %>% 
  mutate_at(tvars,function(x) {x-mean(x,na.rm = T)}) %>% 
  ungroup() %>% 
  #lag
  group_by(region) %>% 
  mutate_at(tvars,list("1"=lag,"2"=function(x) {lag(x,2)})) %>% 
  ungroup() %>% 
  as.data.frame()

#scaled across regions
fdr1=fdr0 %>% 
  #scale
  mutate_at(tvars,scale) %>% 
  #lag
  group_by(region) %>% 
  mutate_at(tvars,list("1"=lag,"2"=function(x) {lag(x,2)})) %>% 
  ungroup() %>% 
  as.data.frame()

#scaled across regions, monthly means removed
fdr1_ds=fdr1 %>% 
  group_by(region,month) %>%
  mutate_at(tvars,list("mm"=function(x) {mean(x,na.rm = T)})) %>% 
  mutate_at(tvars,function(x) {x-mean(x,na.rm = T)}) %>% 
  ungroup() %>% 
  #lag
  group_by(region) %>% 
  mutate_at(tvars,list("1"=lag,"2"=function(x) {lag(x,2)})) %>% 
  ungroup() %>% 
  as.data.frame()
```

Data availability by series and region (prop zeros and prop missing)    

FW: ignore clad, corbic, sside  
W: ignore corbic, sside  
S: ignore potam  

```{r avail}
dataavail=focaldata %>% 
  gather(var, value, 4:length(fvars)) %>% 
  group_by(region, var) %>% 
  summarize(
    propmissing=length(which(is.na(value)))/length(value),
    propzeros=length(which(value==0))/length(which(!is.na(value)))) %>% 
  as.data.frame()

#these variables should not be used (too many zeros)
filter(dataavail,propzeros>0.5 | is.na(propzeros))

#these variables have lots of missing data
filter(dataavail,propmissing>0.45 | is.na(propmissing))
```

### Time series plots

```{r timeseries, fig.width=8, fig.height=13, echo=FALSE, warning=FALSE}
plot1=select(focaldata,region,decyear,varnames) %>% 
  gather(Var,Value,varnames) %>% 
  mutate(Var=factor(Var,levels = varnames),region=factor(region,levels=regionorder))
ggplot(plot1,aes(x=decyear,y=Value)) +
  facet_grid(Var~region, scales="free_y") +
  geom_line() +
  theme_bw() +
  labs(title = "Original Units")

plot2=select(fdr,region,decyear,varnames) %>% 
  gather(Var,Value,varnames) %>% 
  mutate(Var=factor(Var,levels = varnames),region=factor(region,levels=regionorder))
ggplot(plot2,aes(x=decyear,y=Value)) +
  facet_grid(Var~region, scales="free_y") +
  geom_line() +
  theme_bw() +
  labs(title = "Log scaled within region")

plot3=select(fdr_ds,region,decyear,varnames) %>% 
  gather(Var,Value,varnames) %>% 
  mutate(Var=factor(Var,levels = varnames),
         region=factor(region,levels=regionorder))
plotmm=select(fdr_ds,region,decyear,paste0(varnames,"_mm")) %>% 
  gather(Var,Value,paste0(varnames,"_mm")) %>% 
  mutate(Var=gsub("_mm","",Var)) %>% 
  mutate(Var=factor(Var,levels = varnames),
         region=factor(region,levels=regionorder))
ggplot(plot3,aes(x=decyear,y=Value)) +
  facet_grid(Var~region, scales="free_y") +
  geom_line(data=plotmm,color="tomato") +
  geom_line() +
  theme_bw() +
  labs(title = "Seasonal means removed")
```

### Cross-correlation matrices  
(only sig correlations shown... no correction for multiple comparisons)

```{r cc, fig.width=7, fig.height=6, echo=FALSE}
#lag 0
lags0=expand.grid(region=regions,V1=varnames,V2=varnames,Cor=NA,P=NA,
                  stringsAsFactors = F)
for(i in 1:nrow(lags0)) {
  v1=fdr[fdr$region==lags0$region[i],lags0$V1[i]]
  v2=fdr[fdr$region==lags0$region[i],lags0$V2[i]]
  if(!all(is.na(v1)) & !all(is.na(v2))) {
    tccf=cor.test(v1,v2)
    lags0$Cor[i]=tccf$estimate
    lags0$P[i]=tccf$p.value
  }
}
lags0$V1=factor(lags0$V1,levels = varnames)
lags0$V2=factor(lags0$V2,levels = varnames)
lags0$region=factor(lags0$region,levels = regionorder)
ggplot(filter(lags0,P<0.05,V1!=V2), aes(x=V2,y=V1,fill=Cor)) +
  facet_wrap(.~region, nrow=2, scales = "free", drop=F, strip.position = "right",dir = "v") +
  geom_tile() +
  geom_abline(slope=1,intercept = 0) +
  scale_fill_gradient2() +
  theme_bw() +
  labs(x="V2 (t)", y="V1 (t)", title = "lag 0 cross-cor") +
  scale_x_discrete(expand = expand_scale(mult = 0),drop=F) +
  scale_y_discrete(expand = expand_scale(mult = 0),drop=F) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1))

#lag 1
lags1=expand.grid(region=regions,V1=varnames,V2=varnames,Cor=NA,P=NA,
                  stringsAsFactors = F)
for(i in 1:nrow(lags1)) {
  v1=fdr[fdr$region==lags1$region[i],lags1$V1[i]]
  v2=fdr[fdr$region==lags1$region[i],lags1$V2[i]]
  if(!all(is.na(v1)) & !all(is.na(v2))) {
    tccf=cor.test(v1,lag(v2,1))
    lags1$Cor[i]=tccf$estimate
    lags1$P[i]=tccf$p.value
  }
}
lags1$V1=factor(lags1$V1,levels = varnames)
lags1$V2=factor(lags1$V2,levels = varnames)
lags1$region=factor(lags1$region,levels = regionorder)
ggplot(filter(lags1,P<0.05), aes(x=V2,y=V1,fill=Cor)) +
  facet_wrap(.~region, nrow=2, scales = "free", drop=F, strip.position = "right",dir = "v") +
  geom_tile() +
  geom_abline(slope=1,intercept = 0) +
  scale_fill_gradient2() +
  theme_bw() +
  labs(x="V2 (t-1)", y="V1 (t)", title = "lag 1 cross-cor") +
  scale_x_discrete(expand = expand_scale(mult = 0),drop=F) +
  scale_y_discrete(expand = expand_scale(mult = 0),drop=F) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1))

#lag 1 seasonality removed
lags1ds=expand.grid(region=regions,V1=varnames,V2=varnames,Cor=NA,P=NA,
                  stringsAsFactors = F)
for(i in 1:nrow(lags1ds)) {
  v1=fdr_ds[fdr_ds$region==lags1ds$region[i],lags1ds$V1[i]]
  v2=fdr_ds[fdr_ds$region==lags1ds$region[i],lags1ds$V2[i]]
  if(!all(is.na(v1)) & !all(is.na(v2))) {
    tccf=cor.test(v1,lag(v2,1))
    lags1ds$Cor[i]=tccf$estimate
    lags1ds$P[i]=tccf$p.value
  }
}
lags1ds$V1=factor(lags1ds$V1,levels = varnames)
lags1ds$V2=factor(lags1ds$V2,levels = varnames)
lags1ds$region=factor(lags1ds$region,levels = regionorder)
ggplot(filter(lags1ds,P<0.05), aes(x=V2,y=V1,fill=Cor)) +
  facet_wrap(.~region, nrow=2, scales = "free", drop=F, strip.position = "right",dir = "v") +
  geom_tile() +
  geom_abline(slope=1,intercept = 0) +
  scale_fill_gradient2() +
  theme_bw() +
  labs(x="V2 (t-1)", y="V1 (t)",title = "lag 1 cross-cor, monthly means removed") +
  scale_x_discrete(expand = expand_scale(mult = 0),drop=F) +
  scale_y_discrete(expand = expand_scale(mult = 0),drop=F) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1))
```

Other notes:

*Detrended fish indices are NOT correlated in S!*  

Nitrate and ammonia are positively correlated, max at lag 0 all regions.  
Nitrate and dophos are positively correlated, max at lag 0 all regions.  
Ammonia and dophos are positively correlated, lag 0 for FW and S, ammonia lags dphos by 3 months in W and N.  

Chla nitrate neg correlated, lag 0.  
Chla ammonia neg correlated, lag 0.  
Chla dophos relationship unclear.  

High flow 2-4 month prev = high chla

Hcope lags chla by 1, positive, except FW.  
Clad seem to precede chla by 2, positive.  
Amphi relationship unclear, prob bc not eating chla in water column.  
In N and W, chla lags potam, negative. The opposite in W.  

Mysid and hcope postive, lag 0.  
In S and W, hcope lags pcope, negative. 

### Exploratory plots

```{r other, echo=FALSE, eval=FALSE}
#ccf between two variables for all regions
ccfplot=function(v1,v2) {
  par(mfrow=c(2,2),mar=c(3,3,1.5,1))
  for(i in 1:length(regions)) {
    dtemp=filter(fdr,region==regions[i])
    ccf(dtemp$chla, dtemp$flow,lag.max = 6, na.action = na.pass)
    title(paste(regions[i],v1,v2), line=0.5)
  }
}

ccfplot("chla","flow")
ccfplot("hcope","mysid")

#plot multiple variables on top of each other
ggplot(fdr,aes(x=decyear)) + 
  facet_grid(region~.) + 
  geom_line(aes(y=ammonia)) +
  geom_line(aes(y=chla),color="green") +
  geom_line(aes(y=hcope),color="blue")

#plot a single variable
ggplot(fdr,aes(x=decyear, y=flow, color=region)) + geom_line()

#individual bivariate plots
ggplot(focaldata,aes(x=ammonia, y=chla)) + 
  facet_grid(region~.) + geom_point()
ggplot(fdr,aes(x=ammonia_1, y=chla)) + 
  facet_grid(region~.) + geom_point()
ggplot(fdr,aes(x=ammonia, y=dophos, color=chla)) + 
  facet_grid(region~.) + geom_point() + 
  scale_color_viridis_c()
```

### Best predictor of fish

This doesn't consider time lags.

```{r}
l_modelsfw = data.frame(l_model = c("tzoop","tzoop_c",
                                  "hzoop","hzoop+pzoop","pzoop",
                                  "hcope","amphi","pcope",
                                  "tzoop+amphi","hzoop+amphi","pzoop+amphi",
                                  "hcope+pcope","hcope+amphi","pcope+amphi",
                                  "hcope+pcope+amphi"))

l_models = data.frame(l_model = c("tzoop","tzoop_c",
                                  "hzoop","hzoop+pzoop","pzoop",
                                  "mysid","hcope","clad","amphi","pcope",
                                  "tzoop+amphi","hzoop+amphi","pzoop+amphi","pzoop+clad",
                                  "hzoop+mysid","pzoop+clad+hcope",
                                  "hcope+pcope","hcope+amphi","pcope+amphi","hcope+clad",
                                  "amphi+clad","pcope+clad","mysid+amphi",
                                  "hcope+clad+mysid",
                                  "hcope+clad+mysid+pcope",
                                  "hcope+mysid",
                                  "hcope+mysid+pcope",
                                  "hcope+mysid+amphi"))

l_modelss = data.frame(l_model = c("tzoop","tzoop_c",
                                  "hzoop","hzoop+pzoop","pzoop",
                                  "hcope","amphi","pcope","clad",
                                  "tzoop+amphi","hzoop+amphi","pzoop+amphi","pzoop+clad",
                                  "pzoop+clad+hcope",
                                  "hcope+pcope","hcope+amphi","pcope+amphi","hcope+clad",
                                  "amphi+clad","pcope+clad",
                                  "hcope+pcope+amphi","hcope+pcope+clad",
                                  "hcope+pcope+amphi+clad"))

#Midwater Trawl
fwestmods = purrr::map(l_modelsfw$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="Far West" & !is.na(amphi) & !is.na(tzoop))))
names(fwestmods)=l_modelsfw$l_model
bbmle::AICctab(fwestmods,weights=T,delta=T)

westmods = purrr::map(l_models$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="West" & !is.na(amphi) & !is.na(tzoop))))
names(westmods)=l_models$l_model
bbmle::AICctab(westmods,weights=T,delta=T)

northmods = purrr::map(l_models$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="North" & !is.na(amphi) & !is.na(tzoop))))
names(northmods)=l_models$l_model
bbmle::AICctab(northmods,weights=T,delta=T)

southmods = purrr::map(l_modelss$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="South" & !is.na(amphi) & !is.na(tzoop))))
names(southmods)=l_modelss$l_model
bbmle::AICctab(southmods,weights=T,delta=T)
```

Unlike annual model, individual zooplankton groups seems to be better predictors. The sign is not always what you would expect though.

```{r}
l_modelsfw = data.frame(l_model = c("hcope+pcope+amphi",
                                   "hcope+pcope+amphi+marfish_bsmt"))
l_modelsw = data.frame(l_model = c("mysid+amphi",
                                   "mysid+amphi+marfish_bsmt"))
l_modelsn = data.frame(l_model = c("hcope+mysid",
                                   "hcope+mysid+sside"))
l_modelss = data.frame(l_model = c("pzoop+clad+hcope",
                                   "pzoop+clad+hcope+sside"))

#Midwater Trawl
fwestmods = purrr::map(l_modelsfw$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="Far West" & !is.na(amphi))))
names(fwestmods)=l_modelsfw$l_model
bbmle::AICctab(fwestmods,weights=T,delta=T)

westmods = purrr::map(l_modelsw$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="West" & !is.na(amphi))))
names(westmods)=l_modelsw$l_model
bbmle::AICctab(westmods,weights=T,delta=T)

northmods = purrr::map(l_modelsn$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="North" & !is.na(sside))))
names(northmods)=l_modelsn$l_model
bbmle::AICctab(northmods,weights=T,delta=T)

southmods = purrr::map(l_modelss$l_model, ~lm(paste0("estfish_bsmt ~", .x), 
            data = filter(fdr_ds,region=="South" & !is.na(sside))))
names(southmods)=l_modelss$l_model
bbmle::AICctab(southmods,weights=T,delta=T)
```

Marine fishes important in FW but not W. Silversides not important in N and S. 

### Integrated model

This model does not use time lags.

```{r, fig.width=8, fig.height=8}
modFW='chla~potam+flow+temp+secchi
        hcope~chla+potam+flow+temp+secchi
        pcope~chla+potam+flow+temp+secchi
        amphi~chla+potam+flow+temp+secchi
        estfish_bsmt~hcope+pcope+amphi+flow+temp+secchi+marfish_bsmt
        marfish_bsmt~flow
        amphi~~hcope+pcope
        hcope~~pcope
'

modW='chla~potam+flow+temp+secchi
        mysid~chla+potam+flow+temp+secchi
        amphi~chla+potam+flow+temp+secchi
        estfish_bsmt~mysid+amphi+flow+temp+secchi
        amphi~~mysid
'

modN='chla~corbic+flow+temp+secchi
        mysid~chla+corbic+flow+temp+secchi
        hcope~chla+corbic+flow+temp+secchi
        estfish_bsmt~mysid+hcope+flow+temp+secchi
        hcope~~mysid
'

modS='chla~corbic+flow+temp+secchi
        clad~chla+corbic+flow+temp+secchi
        hcope~chla+corbic+flow+temp+secchi
        pzoop~chla+corbic+flow+temp+secchi
        estfish_bsmt~pzoop+clad+hcope+flow+temp+secchi
        hcope~~pzoop+clad
        pzoop~~clad
'

modfitFW=sem(modFW, data=filter(fdr_ds,region=="Far West"))
modfitW=sem(modW, data=filter(fdr_ds,region=="West"))
modfitN=sem(modN, data=filter(fdr_ds,region=="North"))
modfitS=sem(modS, data=filter(fdr_ds,region=="South"))
summary(modfitFW, standardized=T, rsq=T)
summary(modfitW, standardized=T, rsq=T)
summary(modfitN, standardized=T, rsq=T)
summary(modfitS, standardized=T, rsq=T)

#modificationindices(modfitW)

labelsfarwest=createLabels(modfitFW, cnameslag)
labelswest=createLabels(modfitW, cnameslag)
labelsnorth=createLabels(modfitN, cnameslag)
labelssouth=createLabels(modfitS, cnameslag)

#FAR WEST
myLavaanPlot(model=modfitFW, labels=labelsfarwest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#WEST
myLavaanPlot(model=modfitW, labels=labelswest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#NORTH
myLavaanPlot(model=modfitN, labels=labelsnorth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#SOUTH
myLavaanPlot(model=modfitS, labels=labelssouth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
```


### SEM submodels

These models do use time lags.

Note: You get somewhat different answers if you use a multigroup model vs. fit each region individually. I am not entirely sure why, but I think there is some data pooling going on like in hierarchical models. I would have to look into the math. I know that if you use the multigroup framework you can impose constraints like requiring certain parameters to be the same across groups (see [lavaan manual](https://lavaan.ugent.be/tutorial/groups.html)). Obviously, if data on a variable you want to use are missing from one site, you can't use the multigroup option (you will get an error) and have to fit individually.

#### Chla, nutrients, flow

Clams don't seem to have any impact on chla.
Dropping dophos because strongly correlated with ammonia.

```{r, fig.width=8, fig.height=8}
# mod1='chla~ammonia+dophos+chla_1+flow_1
#       ammonia~flow_1+chla_1
#       dophos~flow_1
#       ammonia~~dophos'

modfww='chla~ammonia+chla_1+flow_1+secchi_1+temp_1+potam_1
      ammonia~flow_1+chla_1+secchi_1+temp_1+potam_1'

modns='chla~ammonia+chla_1+flow_1+secchi_1+temp_1+corbic_1
      ammonia~flow_1+chla_1+secchi_1+temp_1+corbic_1'

#you can either do a multigroup model by region
# modfit1=sem(mod1, data=fdr_ds, group = "region")
# summary(modfit1, standardized=T, rsq=T)
# 
# par(mfrow=c(2,2))
# semPaths(modfit1, "std", edge.label.cex = 1, residuals = F,  node.width=3, nCharNodes = 0, intercepts = F, title = T)
# #residuals(modfit1)

#or fit separately for each region 
modfitFW=sem(modfww, data=filter(fdr_ds,region=="Far West"))
modfitW=sem(modfww, data=filter(fdr_ds,region=="West"))
modfitN=sem(modns, data=filter(fdr_ds,region=="North"))
modfitS=sem(modns, data=filter(fdr_ds,region=="South"))
summary(modfitFW, standardized=T, rsq=T)
summary(modfitW, standardized=T, rsq=T)
summary(modfitN, standardized=T, rsq=T)
summary(modfitS, standardized=T, rsq=T)

labelsfarwest=createLabels(modfitFW, cnameslag)
labelswest=createLabels(modfitW, cnameslag)
labelsnorth=createLabels(modfitN, cnameslag)
labelssouth=createLabels(modfitS, cnameslag)

#FAR WEST
myLavaanPlot(model=modfitFW, labels=labelsfarwest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=F, sig=0.05, 
						 width=c("regress","latent","covs"),
						 color=c("regress","latent","covs"))
#WEST
myLavaanPlot(model=modfitW, labels=labelswest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=F, sig=0.05, 
						 width=c("regress","latent","covs"),
						 color=c("regress","latent","covs"))
#NORTH
myLavaanPlot(model=modfitN, labels=labelsnorth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=F, sig=0.05, 
						 width=c("regress","latent","covs"),
						 color=c("regress","latent","covs"))
#SOUTH
myLavaanPlot(model=modfitS, labels=labelssouth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=F, sig=0.05, 
						 width=c("regress","latent","covs"),
						 color=c("regress","latent","covs"))

#residuals(modfitFW)
```

#### Chla, zooplankton

```{r, fig.width=8, fig.height=8}
modFW='chla~chla_1+hcope_1+amphi_1+flow_1+secchi_1+temp_1+potam_1
       amphi~chla_1+amphi_1+flow_1+secchi_1+temp_1+potam_1
       hcope~chla_1+hcope_1+flow_1+pcope_1+secchi_1+temp_1+potam_1
       pcope~chla_1+pcope_1+hcope_1+amphi_1+flow_1+secchi_1+temp_1+potam_1
'
modW='chla~chla_1+hcope_1+flow_1+secchi_1+temp_1+potam_1
       amphi~chla_1+amphi_1+flow_1+secchi_1+temp_1+potam_1
       hcope~chla_1+hcope_1+mysid_1+pcope_1+flow_1+secchi_1+temp_1+potam_1
       pcope~chla_1+pcope_1+hcope_1+mysid_1+flow_1+secchi_1+temp_1+potam_1
       mysid~chla_1+mysid_1+hcope_1+pcope_1+flow_1+secchi_1+temp_1+potam_1
'
modN='chla~chla_1+hcope_1+amphi_1+flow_1+secchi_1+temp_1+corbic_1
       amphi~chla_1+amphi_1+flow_1+secchi_1+temp_1+corbic_1
       hcope~chla_1+hcope_1+flow_1+pcope_1+secchi_1+temp_1+corbic_1
       pcope~chla_1+pcope_1+hcope_1+flow_1+secchi_1+temp_1+corbic_1
       mysid~chla_1+mysid_1+hcope_1+pcope_1+flow_1+secchi_1+temp_1+corbic_1

'
modS='chla~chla_1+hcope_1+clad_1+flow_1+secchi_1+temp_1+corbic_1
       clad~chla_1+clad_1+flow_1+pcope_1+secchi_1+temp_1+corbic_1
       hcope~chla_1+hcope_1+flow_1+pcope_1+secchi_1+temp_1+corbic_1
       amphi~chla_1+amphi_1+flow_1+secchi_1+temp_1+corbic_1
       pcope~chla_1+pcope_1+hcope_1+clad_1+flow_1+secchi_1+temp_1+corbic_1
'
modfitFW=sem(modFW, data=filter(fdr_ds,region=="Far West"))
modfitW=sem(modW, data=filter(fdr_ds,region=="West"))
modfitN=sem(modN, data=filter(fdr_ds,region=="North"))
modfitS=sem(modS, data=filter(fdr_ds,region=="South"))
summary(modfitFW, standardized=T, rsq=T)
summary(modfitW, standardized=T, rsq=T)
summary(modfitN, standardized=T, rsq=T)
summary(modfitS, standardized=T, rsq=T)

#modificationindices(modfitW)

labelsfarwest=createLabels(modfitFW, cnameslag)
labelswest=createLabels(modfitW, cnameslag)
labelsnorth=createLabels(modfitN, cnameslag)
labelssouth=createLabels(modfitS, cnameslag)

#FAR WEST
myLavaanPlot(model=modfitFW, labels=labelsfarwest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#WEST
myLavaanPlot(model=modfitW, labels=labelswest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#NORTH
myLavaanPlot(model=modfitN, labels=labelsnorth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#SOUTH
myLavaanPlot(model=modfitS, labels=labelssouth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))

```

#### Zooplankton, fish

Note that because fish have more missing data, more rows get dropped (compared 'used' and 'total' number of observations to previous summaries).

If you get a warning about variances being negative, that is **definitely** a problem. It seems to happen when you try to create a latent variables, but the variables that go into it don't acutally have a common trend.

```{r, fig.width=8, fig.height=8}
modFW='estfish_bsmt~hcope_1+amphi_1+pcope_1+flow_1+secchi_1+temp_1+marfish_bsmt_1
'
modW='estfish_bsmt~hcope_1+mysid_1+pcope_1+amphi_1+flow_1+secchi_1+temp_1+marfish_bsmt_1
'
modN='estfish_bsmt~hcope_1+mysid_1+pcope_1+amphi_1+flow_1+secchi_1+temp_1+sside_1
'
modS='estfish_bsmt~hcope_1+clad_1+amphi_1+pcope_1+flow_1+secchi_1+temp_1+sside_1
'
modfitFW=sem(modFW, data=filter(fdr_ds,region=="Far West"))
modfitW=sem(modW, data=filter(fdr_ds,region=="West"))
modfitN=sem(modN, data=filter(fdr_ds,region=="North"))
modfitS=sem(modS, data=filter(fdr_ds,region=="South"))
summary(modfitFW, standardized=T, rsq=T)
summary(modfitW, standardized=T, rsq=T)
summary(modfitN, standardized=T, rsq=T)
summary(modfitS, standardized=T, rsq=T)

labelsfarwest=createLabels(modfitFW, cnameslag)
labelswest=createLabels(modfitW, cnameslag)
labelsnorth=createLabels(modfitN, cnameslag)
labelssouth=createLabels(modfitS, cnameslag)

#FAR WEST
myLavaanPlot(model=modfitFW, labels=labelsfarwest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#WEST
myLavaanPlot(model=modfitW, labels=labelswest,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#NORTH
myLavaanPlot(model=modfitN, labels=labelsnorth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
#SOUTH
myLavaanPlot(model=modfitS, labels=labelssouth,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))
```
