---
title: "Annual model - No Regions"
author: "Tanya Rogers"
date: "3/29/2022"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      knit_root_dir = '..',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
library(dplyr)
library(lavaan)
library(DiagrammeR)
library(tidyr)
library(ggplot2)

# For saving SEM diagrams:
library(purrr)
library(DiagrammeRsvg)
library(rsvg)
library(png)
library(grid)
library(ggpubr)
```

### Import data

```{r dataprep}
combined=read.csv("data/annual_averages/annual_data_compiled_noregions.csv")
cnames=read.csv("analysis/column_names.csv", stringsAsFactors = F)
dsub=filter(combined, Year>=1975)
focaldata=dsub[,cnames$Datacolumn]
fvars=cnames$Shortname
colnames(focaldata)=fvars

focaldata = focaldata %>% 
  mutate(tzoop=hcope+clad+mysid+pcope+rotif_m,
         tzoop_e=hcope_e+clad_e+mysid_e+pcope_e+rotif_e,
         hzoop=hcope+clad+rotif_m,
         hzoop_e=hcope_e+clad_e+rotif_e,
         pzoop=mysid+pcope,
         pzoop_e=mysid_e+pcope_e,
         turbid=-secchi) 
fvars=c(fvars,"tzoop","tzoop_e",
        "hzoop","hzoop_e",
        "pzoop","pzoop_e","turbid")
cnames=rbind(cnames,data.frame(Longname=NA,Shortname=c("tzoop","tzoop_e",
                                                       "hzoop","hzoop_e",
                                                       "pzoop","pzoop_e","turbid"),
                               Diagramname=c("total zooplankton",
                                             "total zooplankton\nenergy",
                                             "herbivorous\nzooplankton",
                                             "herbivorous\nzooplankton\nenergy",
                                             "predatory\nzooplankton",
                                             "predatory\nzooplankton\nenergy",
                                             "turbidity"),
                               Datacolumn=NA,Log=c(rep("yes",6),"no"),
                               Color=c("black","black","#ED7D31","#ED7D31","#7030A0",
                                       "#7030A0","#4472C4")))

#focal variables
varnames=c("temp","flow","turbid","chla","hzoop","pzoop","tzoop","amphi","potam","corbic","estfish","estfish_bsmt","estfish_stn")

source("analysis/myLavaanPlot.r")
source("analysis/semDiagramFunctions.r")
```

### Data prep

Log transform, scale

```{r prep}
#log transform
logvars=fvars[cnames$Log=="yes"]
logtrans=function(x) {
  x2=x[which(!is.na(x))]
  if(any(x2==0)) {log(x+min(x2[which(x2>0)],na.rm=T))}
  else {log(x)}
}
focaldatalog = focaldata %>% 
  mutate_at(logvars,logtrans)

#scale data
fd0=focaldatalog
tvars=fvars[-1]

fd=fd0 %>% 
  #lag
  mutate_at(tvars,list("1"=lag)) %>% 
  #scale
  mutate_at(-1,scale) %>% 
  as.data.frame()

#detrended
fd_dtr=fd0 %>% 
  mutate_at(tvars,function(x) { #detrend
    x2=x
    x2[x2==0]=NA
    res=residuals(lm(x2~fd$year))
    out=x
    out[which(!is.na(x2))]=res
    return(out)
  }) %>%
  #lag
  mutate_at(tvars,list("1"=lag)) %>% 
  #scale
  mutate_at(-1,scale)  %>% 
  as.data.frame()
```

### Time series plots

```{r timeseries, fig.width=8, fig.height=8, echo=FALSE, warning=FALSE}
plot1=select(focaldata,year,varnames) %>% 
  gather(Var,Value,varnames) %>% 
  mutate(Var=factor(Var,levels = varnames))
ggplot(plot1,aes(x=year,y=Value)) +
  facet_wrap(Var~., scales="free_y",ncol = 4) +
  geom_line() +
  theme_bw() +
  labs(title = "Original Units")

plot2=select(fd,year,varnames) %>% 
  gather(Var,Value,varnames) %>% 
  mutate(Var=factor(Var,levels = varnames))
ggplot(plot2,aes(x=year,y=Value)) +
  facet_wrap(Var~.,ncol = 4) +
  geom_line() +
  theme_bw() +
  labs(title = "Log scaled within region")

plot3=select(fd_dtr,year,varnames) %>% 
  gather(Var,Value,varnames) %>% 
  mutate(Var=factor(Var,levels = varnames))
ggplot(plot3,aes(x=year,y=Value)) +
  facet_wrap(Var~.,ncol = 4) +
  geom_line() +
  theme_bw() +
  labs(title = "Log scaled detrended")
```

### Other useful plots

Breakdown of total zooplankton biomass.

```{r zoops, echo=FALSE}
plot4=select(focaldata,year,hcope,pcope,clad,mysid,rotif_m) %>% 
  gather(Var,Value,hcope:rotif_m)
#proportion of biomass, on average
plot4 %>% group_by(Var) %>% summarise(Val=mean(Value, na.rm=T)) %>% 
  ungroup %>% mutate(prop=Val/sum(Val))
ggplot(plot4,aes(x=year,y=Value,fill=Var)) +
  geom_area() +
  theme_bw() + scale_fill_brewer(palette = "Dark2") +
  labs(title = "Original units")
```

Similarity of fish indices.

```{r fish, echo=FALSE}
plot5=select(fd,year,estfish,estfish_bsmt,estfish_stn) %>% 
  gather(Var,Value,estfish,estfish_bsmt,estfish_stn) 
ggplot(plot5,aes(x=year,y=Value,color=Var)) +
  geom_line(size=1) +
  theme_bw() + scale_color_brewer(palette = "Dark2") +
  labs(title = "Log scaled")

plot6=select(fd_dtr,year,estfish,estfish_bsmt,estfish_stn) %>% 
  gather(Var,Value,estfish,estfish_bsmt,estfish_stn)
ggplot(plot6,aes(x=year,y=Value,color=Var)) +
  geom_line(size=1) +
  theme_bw() + scale_color_brewer(palette = "Dark2") +
  labs(title = "Log scaled detrended")
```

### Cross-correlation matrices  
(only sig correlations shown... no correction for multiple comparisons)

```{r cc, fig.width=7, fig.height=6, echo=FALSE}
#lag 0
lags0=expand.grid(V1=varnames,V2=varnames,Cor=NA,P=NA,
                  stringsAsFactors = F)
for(i in 1:nrow(lags0)) {
  v1=fd[,lags0$V1[i]]
  v2=fd[,lags0$V2[i]]
  if(!all(is.na(v1)) & !all(is.na(v2))) {
    tccf=cor.test(v1,v2)
    lags0$Cor[i]=tccf$estimate
    lags0$P[i]=tccf$p.value
  }
}
lags0$V1=factor(lags0$V1,levels = varnames)
lags0$V2=factor(lags0$V2,levels = varnames)
ggplot(filter(lags0,P<0.05,V1!=V2), aes(x=V2,y=V1,fill=Cor)) +
  geom_tile() +
  geom_abline(slope=1,intercept = 0) +
  scale_fill_gradient2() +
  theme_bw() +
  labs(x="V2 (t)", y="V1 (t)", title = "lag 0 cross-cor") +
  scale_x_discrete(expand = expand_scale(mult = 0),drop=F) +
  scale_y_discrete(expand = expand_scale(mult = 0),drop=F) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1))

#lag 0, detrended
lags0dtr=expand.grid(V1=varnames,V2=varnames,Cor=NA,P=NA,
                  stringsAsFactors = F)
for(i in 1:nrow(lags0dtr)) {
  v1=fd_dtr[,lags0dtr$V1[i]]
  v2=fd_dtr[,lags0dtr$V2[i]]
  if(!all(is.na(v1)) & !all(is.na(v2))) {
    tccf=cor.test(v1,v2)
    lags0dtr$Cor[i]=tccf$estimate
    lags0dtr$P[i]=tccf$p.value
  }
}
lags0dtr$V1=factor(lags0dtr$V1,levels = varnames)
lags0dtr$V2=factor(lags0dtr$V2,levels = varnames)
ggplot(filter(lags0dtr,P<0.05,V1!=V2), aes(x=V2,y=V1,fill=Cor)) +
  geom_tile() +
  geom_abline(slope=1,intercept = 0) +
  scale_fill_gradient2() +
  theme_bw() +
  labs(x="V2 (t)", y="V1 (t)", title = "lag 0 cross-cor, detrended") +
  scale_x_discrete(expand = expand_scale(mult = 0),drop=F) +
  scale_y_discrete(expand = expand_scale(mult = 0),drop=F) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1))
```

### SEM model

With and without detrending.

```{r}
#1
# model1='zoop=~hcope+clad+mysid
#         fish=~estfish_bsmt+estfish_bsot
#         zoop~chla+potam+flow
#         chla~potam+flow
#         fish~zoop+flow
# '
#2
model1='chla~potam+flow+temp
        hzoop~chla+potam+flow+temp
        pzoop~chla+potam+flow+hzoop+temp
        fish~hzoop+pzoop+flow+turbid+temp+potam
        fish=~estfish+estfish_stn+estfish_bsmt
'

modfit1=sem(model1, data=fd)
modfit1_dtr=sem(model1, data=fd_dtr)
summary(modfit1, standardized=T, rsq=T)
summary(modfit1_dtr, standardized=T, rsq=T)

# residuals(modfit1)
# modificationindices(modfit1, sort=T, maximum.number=20)
```

### Nice plots

#### Without covariances

Original units

```{r, echo=FALSE}
labels1 <- createLabels(modfit1, cnames)

myLavaanPlot(model=modfit1, labels=labels1,
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))

## Updated SEM diagram:
cnames_with_latent <- rbind(cnames,
                            data.frame(Longname="fish", Shortname="fish",
                                       Diagramname="estuarine\nfishes", Datacolumn=NA, Log=NA,
                                       Color="#FF0000"))

plot_modfit1 <- createGraph(fit=modfit1, 
                            reference_df=cnames_with_latent, 
                            model_type="annual",
                            title="Annual no region",
                            manual_port_settings=TRUE)
plot_modfit1
```

Detrended

```{r, echo=FALSE}
myLavaanPlot(model=modfit1_dtr, labels=labels1, 
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=FALSE, sig=0.05, 
						 width=c("regress","latent"),
						 color=c("regress","latent"))

createGraph(fit=modfit1_dtr, 
            reference_df=cnames_with_latent, 
            model_type="annual",
            title="Annual no region, detrended",
            manual_port_settings=TRUE)
```

#### With covariances

Original units

```{r, echo=FALSE}
myLavaanPlot(model=modfit1, labels=labels1, 
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=TRUE, sig=0.05, 
						 width=c("regress","latent","covs"),
						 color=c("regress","latent","covs"))

createGraph(fit=modfit1, 
            reference_df=cnames_with_latent, 
            cov=TRUE,
            model_type="annual",
            title="Annual no region",
            manual_port_settings=TRUE)
```

Detrended

```{r, echo=FALSE}
myLavaanPlot(model=modfit1_dtr, labels=labels1, 
						 node_options=list(shape="box", fontname="Helvetica"), 
						 coefs=TRUE, stand=TRUE, covs=TRUE, sig=0.05, 
						 width=c("regress","latent","covs"),
						 color=c("regress","latent","covs"))

createGraph(fit=modfit1_dtr, 
            reference_df=cnames_with_latent, 
            cov=TRUE,
            model_type="annual",
            title="Annual no region, detrended",
            manual_port_settings=TRUE)
```

*Save updated SEM diagrams*

```{r}
modfit1_grobs <-  map(list(plot_modfit1), ~convert_html_to_grob(.x, 2000))
modfit1_figure <- ggarrange(plotlist=modfit1_grobs) %>%
  annotate_figure(top = text_grob("Annual SEM (whole estuary)",
                                  color = "black",
                                  face = "bold",
                                  size = 8))

ggsave('./fig_output/sem_annual_noregions.png',modfit1_figure, width=3, height=3, dpi=300)
```
