---
title: "Preliminary SEM Analyses 1"
author: "Delta Food Web Team"
date: "10/25/2021"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = '../docs',
      knit_root_dir = '..',
      envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(dplyr)
library(psych) #for pairs.panels, but could use other packages, e.g. GGalley
library(lavaan)
library(semPlot)
```

### Import data

```{r dataprep}
combined=read.csv("data/annual_averages/annual_data_compiled.csv")
cnames=read.csv("analysis/column_names.csv", stringsAsFactors = F)
dsub=filter(combined, Year>=1975)
focaldata=dsub[,cnames$Datacolumn]
fvars=cnames$Shortname
colnames(focaldata)=fvars
```

### Diagnostic plots and data prep

Histograms, log transform, scale

```{r histo}
par(mfrow=c(5,4),mar=c(2,2,2,1))
for(i in 2:length(fvars)) {
  hist(x=focaldata[,i],main=fvars[i])
}
logvars=fvars[cnames$Log=="yes"]
#log transform
logtrans=function(x) {
  x2=x[which(!is.na(x))]
  if(any(x2==0)) {log(x+min(x2[which(x2>0)],na.rm=T))}
  else {log(x)}
}
focaldatalog = focaldata %>% 
  mutate_at(logvars,logtrans)

par(mfrow=c(5,4),mar=c(2,2,2,1))
for(i in 2:length(fvars)) {
  hist(x=focaldatalog[,i],main=fvars[i])
}

#scale data
fd=focaldatalog
fd=fd %>% mutate_at(2:length(fvars),scale) %>% 
  mutate_at(2:length(fvars),list("1"=lag))
```

Time series

```{r timeseries, fig.width=8, fig.height=9}

#original units
par(mfrow=c(5,4),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=focaldata[,1],y=focaldata[,i],main=fvars[i], type="o",cex=0.5)
}

#log scaled
par(mfrow=c(5,4),mar=c(2,2,3,1))
for(i in 2:length(fvars)) {
  plot.default(x=fd[,1],y=fd[,i],main=fvars[i], type="o",cex=0.5)
}
```

Bivariate plots

Flow and salinity are correlated. Flow also shows some correlations with nutrients, temperature. Ammonia and nitrate are correlated.

Striped bass (age-0) and smelt are positively correlated. We might want to try to get data on adult bass, or use some other index for predatory fish, since age-0 bass are probably not eating smelt.

```{r bivariate, fig.width=8, fig.height=8}
#exogenous vars
pairs.panels(select(fd, flow, salinity, secchi, ammonia, nitrate, don, dop, temp),lm = T)
#plankton
pairs.panels(select(fd, chla, hcope, pcal, clad, cyclo, mysid, clams),lm = T)
#fish
pairs.panels(select(fd, smelt, sbass, estfish, marfish),lm = T)

```

### SEM model

Delta smelt

```{r sem1}
model1='chla~temp+flow+secchi+year
        clad~chla+temp+flow+year
        hcope~chla+temp+flow+year
        cyclo~chla+temp+flow+secchi
        clams~chla+hcope+cyclo+pcal+temp+flow+year
        pcal~cyclo+hcope+temp+flow+year
        mysid~cyclo+pcal+hcope+temp+flow+year
        smelt~clad+hcope+pcal+cyclo+mysid+temp+flow+secchi+year
        clams~~0*smelt
        clad~~cyclo
        clad~~hcope
        cyclo~~pcal
        cyclo~~hcope'
modfit1=sem(model1, data=fd)
summary(modfit1, standardized=T, rsq=T)
standardizedSolution(modfit1)
semPaths(modfit1, "std", edge.label.cex = 1, residuals = F)

# residuals(modfit1) #residual covariance matrix
# residuals(modfit1,type="cor") #residual correlation matrix
# modificationIndices(modfit1, standardized=F)
```

Estuarine fishes total

```{r sem2}
model2='chla~temp+flow+secchi+year
        clad~chla+temp+flow+year
        hcope~chla+temp+flow+year
        cyclo~chla+temp+flow+secchi
        clams~chla+hcope+cyclo+pcal+temp+flow+year
        pcal~cyclo+hcope+temp+flow+year
        mysid~cyclo+pcal+hcope+temp+flow+year
        estfish~clad+hcope+pcal+cyclo+mysid+temp+flow+secchi+year
        clams~~0*estfish
        clad~~cyclo
        clad~~hcope
        cyclo~~pcal
        cyclo~~hcope'
modfit2=sem(model2, data=fd)
summary(modfit2, standardized=T, rsq=T)
standardizedSolution(modfit2)
semPaths(modfit2, "std", edge.label.cex = 1, residuals = F)

# residuals(modfit2) #residual covariance matrix
# residuals(modfit2,type="cor") #residual correlation matrix
# modificationIndices(modfit2, standardized=F)
```

Marine fishes total

```{r sem3}
model3='chla~temp+flow+secchi+year
        clad~chla+temp+flow+year
        hcope~chla+temp+flow+year
        cyclo~chla+temp+flow+secchi
        clams~chla+hcope+cyclo+pcal+temp+flow+year
        pcal~cyclo+hcope+temp+flow+year
        mysid~cyclo+pcal+hcope+temp+flow+year
        marfish~clad+hcope+pcal+cyclo+mysid+temp+flow+secchi+year
        clams~~0*marfish
        clad~~cyclo
        clad~~hcope
        cyclo~~pcal
        cyclo~~hcope'
modfit3=sem(model3, data=fd)
summary(modfit3, standardized=T, rsq=T)
standardizedSolution(modfit3)
semPaths(modfit3, "std", edge.label.cex = 1, residuals = F)

# residuals(modfit3) #residual covariance matrix
# residuals(modfit3,type="cor") #residual correlation matrix
# modificationIndices(modfit3, standardized=F)
```
